---
title: "Data_Science_Capstone"
author: "William Terpstra"
date: '2022-06-15'
output: html_document
knit: (function(input_file, encoding) {
    out_dir <- 'docs';
    rmarkdown::render(input_file,
      encoding=encoding,
      output_file=file.path(dirname(input_file), out_dir, 'index.html'))})
---

```{r setup, include=FALSE}
library(Hmisc)
library(dplyr)
library(tidyverse)
library(scales)
library(gridExtra)
library(lmtest)
library(readxl)
library(readr)
library(caret)
library(stringr)
library(leaps) 
library(glmnet)
library(ggplot2)
library(randomForest)
library(pls)
library(MASS)
library(class)
library(gbm)
library(data.table)
library(VIM)
library(tidymodels)
library(themis)
library(kableExtra)
library(kknn)
library(xgboost)
library(vip)
library(missForest)
library(doRNG)
library(doParallel)
library(Boruta)
library(rfUtilities)
library(verification)
library(measures)
library(betacal)
#be more rigorous about necessary dependencies in the future

knitr::opts_chunk$set(echo = TRUE)
```

# Data Science Capstone Project

The purpose of this project is to create data derived estimates of the likelihood and severity of SCA litigation against Mohawk Industries Inc.

```{r Load Data}
#Setting working directory
setwd("C:/Users/Will/Documents/Capstone")
#Loading all the data
ratings <- read_csv("Ratings_Full.csv")
securities <- read_csv("Securities_Full.csv")
SCA_filings <- read_excel("SCA Filings and Settlements.xlsx")
fundamentals <- read_csv("Fundamentals_Full.csv")
stocks <- read.delim("Stocks_DS_tab_delimited.dat", header = TRUE, sep = "\t", dec = ".")
datDict <- read_csv("Data Dictionary.csv")

#Mohawk Industries info:
#its gvkey is 025119
#its GICS industry is 252010
#its GICS group is 2520
```

## Data Cleaning and Feature Engineering

```{r Data Cleaning Functions}
#creating a function to check for NAs
NAcheck <- function(df) {
  names <- c()
  percent_of_missing_values <- c()
  for(i in 1:ncol(df)) { # for-loop over columns in the data frame
    
    #adding the name of each column to a vector
    names <- append(names, colnames(df[i]))
    #adding the amount of missing values of each column to a vector
    percent_of_missing_values <- append(percent_of_missing_values, sum(is.na(df[,i]))/nrow(df))
  }
  #using the two vectors to output a data frame 
  #with the names of columns and their amount of missing values
  data.frame(names, percent_of_missing_values)
}

#creating a function to cull columns that have more missing values than a threshold
NAcull <- function(df, threshold = .20, exception = c("")){
  #threshold is the percentage of tolerable NAs
  #if a column has a higher percentage of NAs then it is culled
  if(threshold > 1){
    stop('threshold must be less than 1, 33% NAs allowed is default')
  }
  drop <- c("")
  
  for(i in 1:ncol(df)) {       
    # for-loop over columns
    #which checks for columns that have over threshold% of data missing,
    #adding the column names to the vector drop
   if(sum(is.na(df[,i]))/nrow(df) > threshold & !(colnames(df[][i]) %in% exception)){
    drop <- append(drop, colnames(df[i]))
    }
  }
  
  #finally the original data frame has all variables that appear in drop
  #removed since they contain more NAs than the threshold%
  output.df <- df[,!(names(df) %in% drop)]
  output.df
}

#function that use quantiles to cap a column's values via 1.5*IQR method
capOutlierIQR <- function(x){
   Q1 <- quantile(x, .25, na.rm = TRUE)
   Q3 <- quantile(x, .75, na.rm = TRUE)
   IQR <- Q3 - Q1
   
   x[ x < Q1 - 1.5*IQR ] <- Q1 - 1.5*IQR
   x[ x > Q3 + 1.5*IQR ] <- Q3 + 1.5*IQR
  
x
}

#function that uses quantiles to clamp values of a column at the 5th and 95th
#percentiles
capOutlier <- function(x){
   Q1 <- quantile(x, .05, na.rm = TRUE)
   Q2 <- quantile(x, .95, na.rm = TRUE)
   
   x[ x < Q1 ] <- Q1
   x[ x > Q2 ] <- Q2
  
x
}

#function that replaces blanks with NA
empty_as_na <- function(x){
    if("factor" %in% class(x)) x <- as.character(x) ## since ifelse wont work with factors
    ifelse(as.character(x)!="", x, NA)
}

#quick googled and copied mode function
#https://www.tutorialspoint.com/r/r_mean_median_mode.htm
getmode <- function(v) {
   uniqv <- unique(v)
   uniqv[which.max(tabulate(match(v, uniqv)))]
}

#method for removing zero variance columns
removeZeroVar <- function(df){
  df[, !sapply(df, function(x) min(x) == max(x))]
}

#method for calculating standard error
std_mean <- function(x) sd(x)/sqrt(length(x))

#methods for mean and sd with NAs removed
#these are used when compressing data sets by summarizing across the 2010-2014 date range
#and the high frequency of NAs is the result of data being collected much more often than reported
na_rm_mean <- function(x) mean(x, na.rm = TRUE)
na_rm_sd <- function(x) sd(x, na.rm = TRUE)

#function for adjusting probabilities after smote up-sampling
#originally from regtools package
#cond probs is the vector of probabilities
#wrongprob1 is the artificially balanced ratio of classes
#trueprob1 is the true ratio of classes
classadjust <- function (condprobs, wrongprob1, trueprob1){
    wrongratio <- (1 - wrongprob1)/wrongprob1
    fratios <- (1/condprobs - 1) * (1/wrongratio)
    trueratios <- (1 - trueprob1)/trueprob1
    1/(1 + trueratios * fratios)
}

#king's transformation; a log transform that deals with negative and 0 values
king_log <- function(x) sign(x)*log(abs(x)+1)
```

```{r Assessing Shared Variables}
#looking to see if any of the data sets share columns
intersect(names(fundamentals), names(ratings))
intersect(names(fundamentals), names(securities))
intersect(names(fundamentals), names(stocks))
intersect(names(ratings), names(securities))
intersect(names(ratings), names(stocks))
intersect(names(securities), names(stocks))

#fixing weird ID column name and data type
names(stocks)[names(stocks) == 'ï..gvkey'] <- 'gvkey'
stocks$gvkey <- as.character(stocks$gvkey)

#seems like the shared variables are mostly demographic information that isn't too useful, 
#however maybe ggroup & gsector could be useful categorical variables
#thus a data frame was created of all 
ggroup_df <- rbind(fundamentals %>% dplyr::select(gvkey, ggroup) %>% distinct(.keep_all = TRUE),
                   ratings %>% dplyr::select(gvkey, ggroup) %>% distinct(.keep_all = TRUE),
                   securities %>% dplyr::select(gvkey, ggroup) %>% distinct(.keep_all = TRUE),
                   stocks %>% dplyr::select(gvkey, ggroup) %>% distinct(.keep_all = TRUE))
#casting ggroup as a categorical variable
ggroup_df$ggroup <- as.factor(ggroup_df$ggroup)

#same process but for gsector
gsector_df <- rbind(fundamentals %>% dplyr::select(gvkey, gsector) %>% distinct(.keep_all = TRUE),
                   ratings %>% dplyr::select(gvkey, gsector) %>% distinct(.keep_all = TRUE),
                   securities %>% dplyr::select(gvkey, gsector) %>% distinct(.keep_all = TRUE),
                   stocks %>% dplyr::select(gvkey, gsector) %>% distinct(.keep_all = TRUE))
gsector_df$gsector <- as.factor(gsector_df$gsector)
```

```{r Adding Primary Key to SCA Dataset}
#The data set of response variables for likelihood and severity modeling, SCA_filings,
#did not have the gvkey variable that uniquely identifies companies and which
#was present in all the data sets with predictor variables
#as a result this code was necessary to append gvkeys to the SCA_filing data set
#by searching for matching company tickers in any of the predictor data sets
#and then appending their corresponding gvkey to the SCA_filings data set


#setting settlement amount to a numeric variable instead of a character
SCA_filings$SettlementAmount <- as.numeric(SCA_filings$SettlementAmount)

#fixing relevant data types for stocks
stocks$tic <- as.character(stocks$tic)

#fixing relevant data types for ratings
ratings$tic <- as.character(ratings$tic)

#fixing relevant data types for securities
securities$tic <- as.character(securities$tic)

#fixing relevant data types for fundamentals
fundamentals$tic <- as.character(fundamentals$tic)

#making a vector of tickers that appear in SCA filings
tic_in_SCA <- SCA_filings %>% filter(FilingYear > 2009 & FilingYear < 2015) %>% dplyr::select("Ticker")

#across all the data I have trying to find occurrences of those tickers
df <- stocks %>% filter(tic %in% tic_in_SCA$Ticker) %>% distinct(gvkey, .keep_all = TRUE) %>% dplyr::select(gvkey, tic, gsector)
df2 <- ratings %>% filter(tic %in% tic_in_SCA$Ticker) %>% distinct(gvkey, .keep_all = TRUE) %>% dplyr::select(gvkey, tic, gsector)
df3 <- securities %>% filter(tic %in% tic_in_SCA$Ticker) %>% distinct(gvkey, .keep_all = TRUE) %>% dplyr::select(gvkey, tic, gsector)
df4 <- fundamentals %>% filter(tic %in% tic_in_SCA$Ticker) %>% distinct(gvkey, .keep_all = TRUE) %>% dplyr::select(gvkey, tic, gsector)

#binding all these occurrences into one data frame
gvkeys_matched_tic <- rbind(df, setdiff(df2, df))
gvkeys_matched_tic <- rbind(gvkeys_matched_tic, setdiff(df3, gvkeys_matched_tic))
gvkeys_matched_tic <- rbind(gvkeys_matched_tic, setdiff(df4, gvkeys_matched_tic)) %>% distinct(.keep_all = TRUE)

#okay now we have all the gvkeys that correspond to 
#companies that appear in the SCA filings data set
skimr::skim(gvkeys_matched_tic)

#for likelihood modeling originally the data was going to be filtered to GICS
#sector 25, the same as Mohawk Industries Inc., but this resulted in a data set
#too small to have a reliable test set for calibration and was ultimately not used
#these variables could be used as filters with the %in% operator for that purpose
gvkeys_matched_tic_notgs25 <- gvkeys_matched_tic %>% filter(gsector != 25) %>% distinct(.keep_all = TRUE)
gvkeys_matched_tic_gs25 <- gvkeys_matched_tic %>% filter(gsector == 25) %>% distinct(.keep_all = TRUE)
```

```{r SCA Fillings Cleaning}
#adding gvkeys to all observations that had their ticker appear in any of
#the predictor data sets
SCA_filings2 <- SCA_filings %>% left_join(gvkeys_matched_tic, 
                                          by = c("Ticker" = "tic"))
#extracting the useful information
SCA_filings2 <- SCA_filings2 %>%
  #making sure it is the same date range as the others
  filter(FilingYear > 2009 & FilingYear < 2015) %>%
  dplyr::select(Ticker, SettlementAmount, gvkey, FilingYear) %>%
  #de-duplication
  distinct(Ticker, .keep_all = TRUE)

#first step in creating the indicator variable for likelihood modeling,
#assigning Sued = Yes to every company that appears in the SCA filings data set
SCA_filings2 <- SCA_filings2 %>% mutate(Sued = "Yes")

#one last look to verify it is well formatted
#skimr::skim(SCA_filings2)
```

```{r Ratings Cleaning}
#Seems the data is rife with missing values, to the extent it cannot be used
#but credit ratings seem like a variable that would be very important for predicting
#shareholder litigation and furthermore they are the only unique variable present 
#in this data set, consequently I will proceed with the inclusion of 
#splticrm for this project, splticrm is the long term issuer credit rating, 
#and the credit rating with the lowest amount of NAs
NAcheck(ratings)

#assessing date range
ratings$datadate <- strptime(ratings$datadate, format = "%m/%d/%Y")
summary(ratings$datadate)

#filtering out all the unnecessary variables and only including companies in the same
#GIC sector as Mohawk Industries Inc. (Code 252010, Household Durables)
ratings2 <- ratings %>% 
  #filter(gsector == 25| gvkey %in% gvkeys_matched_tic$gvkey) %>% #old code to filter by gsector
  dplyr::select(gvkey, conml, splticrm, datadate, tic, gsector)

#de-duplication
ratings2 <-  ratings2 %>% distinct(.keep_all = TRUE)

#Once again checking NAs to assess the data quality with this portion of the dataset
#actually a slight improvement
NAcheck(ratings2)

#looking at the distribution of NAs per company
ratings2 %>% group_by(gvkey) %>% summarise(na = sum(is.na(splticrm))) %>% ggplot(aes(x=na)) + geom_histogram()

#mutating credit ratings to a numeric value, since it is a ordinal categorical variable
ratings2 <- ratings2 %>% 
  mutate(splticrm = case_when(splticrm == "D" ~ "1",
                        splticrm == "C-"  ~ "2",
                        splticrm == "CC"  ~ "3",
                        splticrm == "CCC-"  ~ "4",
                        splticrm == "CCC+"  ~ "5",
                        splticrm == "B-"  ~ "6",
                        splticrm == "B"  ~ "7",
                        splticrm == "B+"  ~ "8",
                        splticrm == "BB-"  ~ "9",
                        splticrm == "BB"  ~ "10",
                        splticrm == "BB+"  ~ "11",
                        splticrm == "BBB-"  ~ "12",
                        splticrm == "BBB"  ~ "13",
                        splticrm == "BBB+"  ~ "14",
                        splticrm == "A-"  ~ "15",
                        splticrm == "A"  ~ "16",
                        splticrm == "A+"  ~ "17",
                        splticrm == "AA-"  ~ "18",
                        splticrm == "AA"  ~ "19",
                        splticrm == "AA+"  ~ "20",
                        splticrm == "AAA"  ~ "21",
                        TRUE  ~ "NA"))

#making sure R recognizes the column data is now num not chr
ratings2 <- ratings2 %>% mutate(splticrm = as.numeric(splticrm))
summary(ratings2$splticrm)

#Since the data occurs over a time span and time series data is tricky
#for the models I know how to use (since it is generally auto correlated)
#I address this by condensing credit rating to the mean and standard deviation
#across the given time span
#all data is technically time series data anyways
ratings2 <- ratings2 %>% group_by(gvkey) %>% mutate(splticrm_avg = mean(splticrm, na.rm = TRUE)) %>% 
  mutate(splticrm_sd = sd(splticrm, na.rm = TRUE))

#selecting only teh variables of interest
ratings2 <- ratings2 %>% dplyr::select(gvkey, splticrm_avg, splticrm_sd) %>% distinct(gvkey, .keep_all = TRUE)

#assessing NAs
NAcheck(ratings2)

#seems everything worked, and the data is ready to be joined to other data sets
#skimr::skim(ratings2)
```

```{r Securities Cleaning}
#selecting the variables of interest, excluding ones like address, phone number, etc.
securities2 <- securities %>%
  #filter(gsector == 25 | gvkey %in% gvkeys_matched_tic$gvkey) %>% #old code to filter by gsector
  dplyr::select(-c(datadate, iid, conm, isalrt, primiss, spgim, 
                   spiim, spmim, curcddvm, sph100, sphcusip, sphiid, 
                   sphmid, sphname, sphsec, sphtic,
                   cyear, mkvalincl, exchg, tpci, city, cmth, conml, costat,
                   ggroup, gind, gsector, gsubind, loc, naics, sic, state))

#correcting a categorical variable's data type
securities2$curcdm <- as.factor(securities2$curcdm)

#assessing data completion
NAcheck(securities2) %>% arrange(percent_of_missing_values)

#assessing data date range
#seems like the standard 2010-2014 range
#securities2$datadate <- strptime(securities2$datadate, format = "%m/%d/%Y")
#summary(securities2$datadate)

#de-duplication
securities2 <-  securities2 %>% distinct(.keep_all = TRUE)

#checking how NAs are distributed between companies via a histogram
securities2 %>% group_by(gvkey) %>% summarise(na = sum(is.na(dvrate))) %>% ggplot(aes(x=na)) + geom_histogram()
securities2 %>% ungroup()

#Now to address the fact this is data over a date range, I'm going to have to
#summarize the data, essentially use functions to condense it into a single
#observation per company.
#Numeric variables are summarized as their mean and standard deviation
#Nominal categorical variables are summarized by their mode
#Ordinal categorical variables are transformed into numeric variables then
#also summarized by their mean and standard deviation
securities2 <- securities2 %>% group_by(gvkey) %>%
   summarise_if(is.numeric, list(avg = na_rm_mean, sd = na_rm_sd))

#checking variables with a suspicious amount of 0s
lapply(securities2, function(x){ length(which(x==0))/length(x)})

#removing variables with too many missing values
 securities2 <- NAcull(securities2)
 NAcheck(securities2)

#looking at univariate distributions
ggplot(gather(securities2 %>% dplyr::select(where(is.numeric))), aes(value)) + 
    geom_histogram(fill="#69b3a2", color="#e9ecef", alpha=0.9, bins = 20) + 
    facet_wrap(~key, scales = 'free_x') +
    labs(title = "Securities Univariate Distributions")

#looking at log transforms of univariate distributions
ggplot(gather(securities2 %>% dplyr::select(where(is.numeric))), aes(log(value))) + 
    geom_histogram(fill="#69b3a2", color="#e9ecef", alpha=0.9, bins = 20) + 
    facet_wrap(~key, scales = 'free_x') +
    labs(title = "Securities Univariate Distributions Log Transformed")

#capping outliers
idx <- sapply(securities2, class)== "numeric"
securities2[, idx] <- lapply(securities2[, idx], capOutlier)

#adding log transformed variables that seemed justified
#due to less skew but mainly the salience of relative changes over absolute
securities2 <- securities2 %>% 
  mutate_at(.vars = c("cshtrm_avg", "cshtrm_sd", "prchm_avg", "prchm_sd", "trt1m_avg", "trt1m_sd"), 
            .funs = list(log = king_log))

#adding new features
securities2 <- securities2 %>% 
  #adding a new feature, share turnover, monthly
  mutate(shrtrn_m_avg = cshtrm_avg/cshom_avg) %>%
  #adding a new feature, highest and lowest share price difference monthly
  mutate(prchl_m_diff = prchm_avg - prclm_avg) %>%
  #replacing NaNs with NAs
  mutate_all(function(x) ifelse(is.nan(x), NA, x))
  
#assessing missing values
NAcheck(securities2) %>% arrange(percent_of_missing_values)

#verifying the data is in the format I want
#skimr::skim(securities2)
```

```{r Stocks Cleaning}
#only care about gcis sector
#stocks <- stocks %>% filter(gsector == 25 | gvkey %in% gvkeys_matched_tic$gvkey) #old code to filter by gsector

#assessing date range
#seems like the standard 2010-2014 range
#stocks$datadate <- strptime(stocks$datadate, format = "%m/%d/%Y")
#summary(stocks2$datadate)

#removing irrelevant columns like phone, company name, address
stocks2 <- stocks[,c(1, 8:11, 13, 14, 24, 26:29, 31:37, 73)]

#setting data misidentified as numeric to factors
stocks2$stko <- as.factor(stocks2$stko)
stocks2$prcstd <- as.factor(stocks2$prcstd)
stocks2$exchg <- as.factor(stocks2$exchg)

#de-duplication
stocks2 <-  stocks2 %>% distinct(.keep_all = TRUE)

#looking at the frequency of zeroes in the data set
#to weed out any imputation errors/irregularities
lapply(stocks2, function(x){ length(which(x==0))/length(x)})

#seems like a potentially suspicious amount of 0s are in cshtrd based on the
#log distribution, but ultimately it seems plausible 
#that some stocks just don't have shares traded often
ggplot(stocks2, aes(x=  sign(stocks2$cshtrd)*log(abs(stocks2$cshtrd)+1))) + 
  geom_histogram(fill="#69b3a2", color="#e9ecef", alpha=0.9) +
  labs(title = "Transformation Identifying Potential NAs", x = "Log(cshtrd)", y = "Count")

#once again condensing the numeric data so there is one observation per company (gvkey)
stocks_num <- stocks2 %>% group_by(gvkey) %>%
   summarise_if(is.numeric, list(avg = na_rm_mean, sd = na_rm_sd))

#looking at missing values
NAcheck(stocks_num) %>% arrange(percent_of_missing_values)

#removing columns with excessive missing values after each observation corresponds to a company
 stocks_num <- stocks_num %>% ungroup() %>% NAcull(threshold = .4)
 NAcheck(stocks_num)

#looking at univariate distributions
ggplot(gather(stocks_num %>% dplyr::select(-gvkey)), aes(value)) + 
    geom_histogram(fill="#69b3a2", color="#e9ecef", alpha=0.9, bins = 15) + 
    facet_wrap(~key, scales = 'free_x') +  
    labs(title = "Stocks Univariate Distributions")

#looking at log transforms of univariate distributions
ggplot(gather(stocks_num %>% dplyr::select(-gvkey)), aes(log(value))) + 
    geom_histogram(fill="#69b3a2", color="#e9ecef", alpha=0.9, bins = 15) + 
    facet_wrap(~key, scales = 'free_x') + 
    labs(title = "Stocks Univariate Distributions Log Transformed")

#capping outliers
idx <- sapply(stocks_num, class)== "numeric"
stocks_num[, idx] <- lapply(stocks_num[, idx], capOutlier)

#adding log transformed variables that seem justified
#due to less skew but mainly due to the salience of relative changes over absolute
stocks_num <- stocks_num %>% 
  mutate_at(.vars = c("cshoc_avg", "cshoc_sd", 
                      "cshtrd_avg", "cshtrd_sd", 
                      "prchd_avg", "prchd_sd",
                      "trfd_avg", "trfd_sd"), 
            .funs = list(log = king_log)) %>% 
  #adding a new feature, share turnover daily
  mutate(shrtrn_d_avg = cshtrd_avg/cshoc_avg)

#replacing NaNs with NAs
stocks_num <- stocks_num %>% mutate_all(function(x) ifelse(is.nan(x), NA, x))

#also collapsing the date range for factors by taking the mode 
#of each factor variable
stocks_fac <- stocks2 %>% 
  group_by(gvkey) %>%
  summarise_if(is.factor, getmode)
stocks_fac %>% ungroup()

#combining the collapsed numeric and factor data sets
stocks2 <- stocks_num %>% left_join(stocks_fac, by = c("gvkey"))

#assessing the amount of missing values one last time
NAcheck(stocks2) %>% arrange(percent_of_missing_values)

#verifying the data is in the format I want
#skimr::skim(stocks2)
```

```{r Fundamentals Cleaning}
#first need to limit data to one format so there isn't repeats
fundamentals2 <- fundamentals %>% 
  #filter(gsector == 25 | gvkey %in% gvkeys_matched_tic$gvkey) %>% #old code to filter by gsector
  filter(datafmt == "STD")

#assessing date range
#in sync with all the others
# fundamentals2$datadate <- strptime(fundamentals2$datadate, format = "%m/%d/%Y")
# summary(fundamentals2$datadate)

#de-duplication
fundamentals2 <-  fundamentals2 %>% distinct(.keep_all = TRUE)

#I inner join a csv of the data dictionary
fundamentals2 %>% NAcheck() %>% inner_join(datDict, by = c("names" = "Mnemonic")) %>% arrange(percent_of_missing_values)

#now removing variables that aren't useful in the model like company name, address
#and business description
fundamentals2 <- fundamentals2 %>% dplyr::select(-any_of(c("fyear", "conm", "acctstd", "fyr", "src", 
                                                "apdedate", "fdate", "exchg", "costat", 
                                                "naicsh", "sich", "add1", "addzip", "busdesc", "city", "conml",
                                                "ein", "fyrc", "gsubind", "incorp", "gsector", 
                                                "ggroup", "loc", "naics", "phone", "priusa", "sic", 
                                                "weburl", "auopic")))

#casting the categorical variables of interest to factors
factors <- c("stko", "idbflag", "auop", "ceoso", "cfoso")
fundamentals2[,factors] <- lapply(fundamentals2[,factors] , factor)

#summarizing the factor variables
fund_factors <- fundamentals2 %>% 
  group_by(gvkey) %>%
  summarise_if(is.factor, getmode)
str(fund_factors)

#summarizing the numeric variables
fund_num <- fundamentals2 %>%
  group_by(gvkey) %>%
  summarise_if(is.numeric, list(avg = na_rm_mean, sd = na_rm_sd))

#adding some new features
#Price-earnings ratio
fund_num <- fund_num %>% mutate(PE_ratio_avg = epspx_avg/prcc_c_avg)
#fiscal year difference between highest and lowest share price
fund_num <- fund_num %>% mutate(pchl_fy_diff = prcl_f_avg - prch_f_avg)
#share turnover
fund_num <- fund_num %>% mutate(shrtrn_f_avg = cshtr_f_avg/csho_avg)
#replacing any NaNs with NAs
fund_num <- fund_num %>% mutate_all(function(x) ifelse(is.nan(x), NA, x))

#joining the summarized numeric and factor variables
fundamentals2 <- fund_factors %>% left_join(fund_num, by = c("gvkey"))

#capping outliers
idx <- sapply(fundamentals2, class)== "numeric"
fundamentals2[, idx] <- lapply(fundamentals2[, idx], capOutlier)
```

## Likelihood Modeling

#### Predicting the likelihood of SCA litigation for Mohawk Industries Inc.

The best model was selected using k-folds cross validation on the training set. It was then calibrated using k-folds cross validation and had its final performance assessed on a held out test set.

#### Final data set creation with missing value imputation using MissForest.
MissForest imputation was selected as the missing value imputation method because it is a robust imputation method that is nonparametric, can handle numeric and categorical variables, is computationally inexpensive, can handle high dimensionality, and provides error estimates using OOB error.

```{r Final Dataset Cleaning and Imputation for Likelihood}
#creating the final likelihood data set
#inner joining the most information dense and rigorously reported data sets
#left joining the rest
Data <- securities2 %>% inner_join(fundamentals2, by = c("gvkey"))
Data <- Data %>% left_join(stocks2, by = c("gvkey"))
Data <- Data %>% left_join(ratings2, by = c("gvkey"))
Data <- Data %>% left_join(ggroup_df %>% distinct(.keep_all = TRUE), by = c("gvkey"))

#Assessing missing values
NAcheck(Data) %>% arrange(percent_of_missing_values)

#Adding the response variables
Data <- Data %>% left_join(SCA_filings2, by = c("gvkey"))

#Removing ticker since it was only used for joining this data frame
Data <- Data %>%
  #old code for removing observations that weren't in gsector 25
  #filter(!gvkey %in% gvkeys_matched_tic_notgs25$gvkey) %>%
  dplyr::select(-c(Ticker, FilingYear))

#Adding the "No"s for the indicator variable
Data$Sued <- ifelse(is.na(Data$Sued), "No", "Yes")
#ordering the factors which is important for yardstick functions
Data$Sued <- factor(Data$Sued, levels =c("Yes", "No"))
#making sure everything worked correctly
summary(Data$Sued)
#checking NAs, quite a lot
NAcheck(Data) %>% arrange(percent_of_missing_values)

#removing NAs
Data <- NAcull(Data)

#visualizing missing data to help with MAR assessment
#aggr_plot <- aggr(Data, col=c('navyblue','red'), numbers=TRUE, sortVars=TRUE, labels=names(data), cex.axis=.7, gap=3, ylab=c("Histogram of missing #data","Pattern"))

#running in parallel to speed up performance
#also setting a seed for reproducible results in parallel
doParallel::registerDoParallel()
set.seed(123, kind = "L'Ecuyer-CMRG")

#Saving settlement data so it can be removed and reattached
#so it isn't accidentally imputed
Data_Settle <- Data %>% dplyr::select(gvkey)
Data <- Data %>% dplyr::select(-gvkey)
Data <- as.data.frame(Data)

#performing missForest imputation
#https://pubmed.ncbi.nlm.nih.gov/22039212/
#a good non parametric NA solution
Data_imputed <- missForest(Data, 
                           parallelize = "forests", 
                           verbose = TRUE, 
                           maxiter = 1, 
                           ntree = 50)
#assessing imputation error
Data_imputed$OOBerror

#checking to make sure that the data is properly imputed
Best_impute <- Data_imputed$ximp
NAcheck(Best_impute)

#reattaching settlement amount data after imputation
Data <- cbind(Data_Settle, Best_impute)

#calculating Sued class imbalance
sum(Data$Sued == "Yes")/sum(Data$Sued == "No")

#removing Mohawk Industries Observation to predict later
Mohawk_Obs <- Data %>% filter(gvkey == "025119")

#making sure the Mohawk Industries observation is properly saved
Mohawk_Obs

#removing Mohawk Industries Observation so it isn't used in either
#the training or testing sets
Data <- Data %>% filter(gvkey != "025119")
```

#### The Boruta Algorithm was used for feature selection.

```{r Boruta Feature Selection for Likelihood}
#to select features from the >200 predictor model
#ideally this would be nested within each cross validation fold of the model
#building procedure but that was not computationally viable

set.seed(111)

#for the universal part of the feature selection process for likelihood models 
#I use Boruta, a wrapper method that selects all relevant variables (not minimal)
boruta <- Boruta(Sued ~ ., data = Data[,-c(1,2)], doTrace = 2, maxRuns = 60)
print(boruta)

#the default plot of boruta, a bit too dense to be that informative besides
#the fact it is working
plot(boruta, las = 2, cex.axis = 0.7)
attStats(boruta) %>% arrange(desc(meanImp))

#tentative rough fix deals with attributes that are still tentative features
#after maxRuns is reached
final_boruta <- TentativeRoughFix(boruta)
features_selected <- getSelectedAttributes(boruta)
features_selected <- c(features_selected, "Sued", "gvkey")
features_selected
#plotting the top 10% of variables by mean importance in regards to being sued
attStats(boruta) %>%
  mutate(Variable = rownames(attStats(boruta))) %>%
  dplyr::filter(meanImp > quantile(attStats(boruta)$meanImp, .50)) %>%
  mutate(Boruta_Mean_Importance = meanImp,
         Variable = fct_reorder(Variable, meanImp)) %>%
  ggplot(aes(x = Boruta_Mean_Importance, y = Variable)) +
  geom_col(fill = "paleturquoise3") +
  labs(title = "Top 5% Most Important Variable for Sued as Response", y = NULL)
```

#### A tidymodels recipe was used to remove all zero variance variables, normalize the data, and one hot encode all categorical variables.

```{r Likelihood Generic Preprocessing}
#Generic tidymodels recipe for pre-processing likelihood data

#assessing correlations for potential interactions
#corrplot::corrplot(cor(as.matrix(Data[features_selected] %>% dplyr::select_if(is.numeric))))

#formula for identifying predictors
Data_rec <- recipe(Sued ~ ., data = Data[,features_selected]) %>%
  #updating the variable gvkey to a non-predictor role
  update_role(gvkey, new_role = "ID") %>%
  #normalization so feature scales are commensurate
  step_normalize(all_numeric(), -gvkey) %>%
  #turning nominal factors into dummy variables 
  #so they are compatible with certain modeling functions
  step_dummy(all_nominal(), -Sued, -gvkey) %>%
  #smote to address class imbalance but will require calibration
  #on second thought better to not disturb the underlying probabilities
  #step_smote(Sued, over_ratio = 1) %>%
  #important to note Boruta does not address colinearity
  step_corr(all_numeric(), threshold = 0.9, method = "spearman") %>%
  #removing all zero variance variables
  step_zv(all_predictors())

#code to assess if the recipe is working properly
# test_df <- Data_rec %>% # use the recipe object
#   prep() %>% # perform the recipe on training data
#   juice() # extract only the pre-processed data frame 
```

```{r Train Test Split}
#creating the test and training split of data
set.seed(111)
#boruta selected var set
Data_split <- Data[,features_selected] %>% initial_split(strata = Sued, prop = 75/100)
Data_train <- training(Data_split)
Data_test <- testing(Data_split)

#creating folds for 5 fold stratified cross validation
Data_folds <- vfold_cv(Data_train, v = 5, strata = Sued)

#creating resamples to be used for calibration
Data_cal_folds <- vfold_cv(Data_train, strata = Sued, v = 5)
```

#### Log-loss was used as the model evaluation metric since ideally we want a model that not only labels well but produces well calibrated probabilities.

```{r Tidymodels Generic Work Flow}
#generic work flow
Data_wf <- workflow() %>%
  add_recipe(Data_rec)

#for extracting coefficients
get_model <- function(x) {
  pull_workflow_fit(x) %>% tidy()
}

#setting metrics to collect
metrics <- metric_set(yardstick::accuracy, yardstick::mn_log_loss, yardstick::specificity, yardstick::precision, yardstick::recall, yardstick::roc_auc)
```

#### Performing the flat stratified k-fold cross validation procedure for model selection. 

```{r lr_model}
#logistic regression model specifications
#mixture 1 indicates this is a pure lasso model
#thus variable selection is performed
lr_spec <- logistic_reg(penalty = tune(), mixture = 1) %>% 
  #important to set family to binomial since this is a classification problem
  set_engine("glmnet", family="binomial")

#grid for tuning the penalty of the logistic regression model
lr_grid <- tibble(penalty = 10^seq(3, -4, length.out = 40))

#tuning the model using k-folds cross validation
lr_tune_results <- Data_wf %>%
  add_model(lr_spec) %>% 
  tune_grid(Data_folds,
            grid = lr_grid,
            control = control_grid(save_pred = TRUE),
            metrics = metrics)

#creating a df containing the optimal penalty to highlight in a graph
highlight_df <- lr_tune_results %>% 
  collect_metrics() %>% 
  filter(.metric == "mn_log_loss") %>%
  filter(mean == min(mean))

#plotting the optimal lasso penalty by lowest log loss
  lr_tune_results %>% 
  collect_metrics() %>% 
  filter(.metric == "mn_log_loss") %>%
  ggplot(aes(x = penalty, y = mean)) + 
  geom_point(color = "#7CAE00") + 
  ylab("mn_log_loss") +
  scale_x_log10(labels = scales::label_number()) +
  geom_line(color = "#7CAE00") +
  geom_point(data=highlight_df, size = 2, color = "red2") +
  geom_vline(xintercept= highlight_df$penalty, 
             linetype="dashed", 
             color = "red",
             size = .5) +
  labs(title = "Optimal Penalty", y = "Log Loss", x = "Penalty" )

#selecting the best model
#since we care about the underlying probability rather than label
#I use log loss to determine the best model
lr_best <- lr_tune_results %>%
  select_best("mn_log_loss")

#collecting the predictions of the best model
#and mutating the column to identify the kind of model
lr_train_best_pred <- lr_tune_results %>% 
  collect_predictions(parameters = lr_best) %>%
  mutate(model = "Logistic Regression")

#plotting the roc_curve for the best lr model
lr_train_best_pred %>% roc_curve(Sued, .pred_Yes) %>% autoplot()

#threshold assessment for the confusion matrix
#calculating the ideal threshold using the Fbeta measure
lr_best_thresh_dat <- lr_train_best_pred %>%
  filter(.config == lr_best$.config) %>%
  pr_curve(Sued, .pred_Yes) %>%
  #F1-measure 
  mutate("F1 Measure" = ((1 + 1^2) * precision * recall) / ((1^2) * precision + recall)) %>%
  #F2-measure emphasizes minimizing false negatives
  mutate("F2 Measure" = ((1 + 2^2) * precision * recall) / ((2^2) * precision + recall)) %>%
  #F3-measure emphasizes minimizing false negatives even more
  mutate("F3 Measure" = ((1 + 2^3) * precision * recall) / ((2^3) * precision + recall)) %>%
  pivot_longer(!c(.threshold, recall, precision),
               names_to = "Fbeta_Measure", 
               values_to = "value")

#calculating ideal threshold using balanced accuracy
lr_best_thresh_dat2 <- lr_train_best_pred %>%
  filter(.config == lr_best$.config) %>%
  roc_curve(Sued, .pred_Yes) %>%
  mutate("Bal_Acc" = (specificity + sensitivity)/2)

#saving the best threshold using balanced accuracy
lr_best_thresh_BA <- lr_best_thresh_dat2 %>% filter(Bal_Acc == max(Bal_Acc, na.rm = TRUE))
lr_best_thresh_BA <- lr_best_thresh_BA$.threshold

#saving the ideal threshold using the F2 measure
lr_best_thresh <- lr_best_thresh_dat %>% filter(Fbeta_Measure == "F2 Measure") %>% filter(value >= max(value, na.rm = TRUE))
lr_best_thresh <- lr_best_thresh$.threshold

#plot of thresholds
lr_best_thresh_dat %>%
  ggplot(aes(x = .threshold, y = value, color = Fbeta_Measure)) +
  geom_path(show.legend = TRUE, alpha = 0.6, size = .75) +
  coord_equal() + 
  geom_vline(xintercept= max(lr_best_thresh), 
             linetype="dashed", 
             color = "red",
             size = .5) +
  labs(title = "Optimal Classifier Threshold for Linear Regression", y = "Value", x = "Threshold" )

#the confusion matrix of the best logistic regression model
lr_tune_results %>% 
  collect_predictions() %>%
  filter(.config == lr_best$.config) %>%
  #custom threshold
  mutate(.pred_class = ifelse(.pred_Yes > lr_best_thresh, "Yes", "No")) %>%
  conf_mat(Sued, .pred_class)

#the metrics of the best logistic regression model
lr_tune_results %>% 
  collect_metrics() %>%
  filter(.config == lr_best$.config)

#finalizing logistic regression work flow with the best hyper parameters
lr_wf <- Data_wf %>% add_model(lr_spec)
final_lr <- finalize_workflow(lr_wf, lr_best)

#performing the last fit
final_lr_test_results <- last_fit(final_lr, Data_split, metrics = metrics)

#collecting the final predictions
collect_metrics(final_lr_test_results)
```

```{r rf_model}
set.seed(345)

#random forest tuneable model specifications
rf_spec <- rand_forest(mtry = tune(), min_n = tune()) %>%
  #mtry is the number of variables randomly sampled as candidates at each split
  #min_n is the number of trees to build (from which classification is decided)
  set_engine("randomForest", num.threads = parallel::detectCores()) %>% 
  set_mode("classification")

#tuning procedure using k-folds cross validation
rf_tune_results <- Data_wf %>%
  add_model(rf_spec) %>% 
  tune_grid(Data_folds,
            grid = 25, #25 candidate models
            control = control_grid(save_pred = TRUE),
            metrics = metrics)

#again since we care more about the underlying probabilities
#rather than labels I go with mn_log_loss
rf_tune_results %>% 
  show_best(metric = "mn_log_loss")

#visualizing tune results
autoplot(rf_tune_results)

#selecting the best model by optimal log loss
rf_best <- rf_tune_results %>% 
  select_best("mn_log_loss")

#collecting the predictions of the best model
#and mutating the column to identify the kind of model
rf_train_best_pred <- rf_tune_results %>% 
  collect_predictions(parameters = rf_best) %>%
  mutate(model = "Random Forest")

#plotting the roc_curve for the best rf model
rf_train_best_pred %>% 
  filter(.config == rf_best$.config) %>%
  roc_curve(Sued, .pred_Yes) %>% 
  ggplot(aes(1 - specificity, sensitivity)) +
  geom_abline(lty = 2, color = "gray80", size = 1) +
  geom_path(show.legend = FALSE, alpha = 0.6, size = .75) +
  coord_equal()

#threshold assessment for the confusion matrix
#calculating the ideal threshold using the Fbeta measure
rf_best_thresh_dat <- rf_train_best_pred %>%
  filter(.config == rf_best$.config) %>%
  pr_curve(Sued, .pred_Yes) %>%
  #F1-measure 
  mutate("F1 Measure" = ((1 + 1^2) * precision * recall) / ((1^2) * precision + recall)) %>%
  #F2-measure emphasizes minimizing false negatives
  mutate("F2 Measure" = ((1 + 2^2) * precision * recall) / ((2^2) * precision + recall)) %>%
  #F3-measure emphasizes minimizing false negatives even more
  mutate("F3 Measure" = ((1 + 2^3) * precision * recall) / ((2^3) * precision + recall)) %>%
  pivot_longer(!c(.threshold, recall, precision),
               names_to = "Fbeta_Measure", 
               values_to = "value") 

#calculating ideal threshold using balanced accuracy
rf_best_thresh_dat2 <- rf_train_best_pred %>%
  filter(.config == rf_best$.config) %>%
  roc_curve(Sued, .pred_Yes) %>%
  mutate("Bal_Acc" = (specificity + sensitivity)/2)

#saving the best threshold using balanced accuracy
rf_best_thresh_BA <- rf_best_thresh_dat2 %>% filter(Bal_Acc == max(Bal_Acc, na.rm = TRUE))
rf_best_thresh_BA <- rf_best_thresh_BA$.threshold

#saving the ideal threshold using F2 measure
rf_best_thresh <- rf_best_thresh_dat %>% filter(Fbeta_Measure == "F2 Measure") %>% filter(value == max(value, na.rm = TRUE))
rf_best_thresh <- rf_best_thresh$.threshold

#plot of thresholds
rf_best_thresh_dat %>%
  ggplot(aes(x = .threshold, y = value, color = Fbeta_Measure)) +
  geom_path(show.legend = TRUE, alpha = 0.6, size = .75) +
  coord_equal() + 
  geom_vline(xintercept= max(rf_best_thresh), 
             linetype="dashed", 
             color = "red",
             size = .5) +
  labs(title = "Optimal Classifier Threshold for Random Forest", y = "Value", x = "Threshold" )

#the confusion matrix of the best rf model on the train set
rf_tune_results %>% 
  collect_predictions() %>%
  filter(.config == rf_best$.config) %>%
  #custom threshold
  mutate(.pred_class = ifelse(.pred_Yes > rf_best_thresh, "Yes", "No")) %>%
  conf_mat(Sued, .pred_class)
  
#the metrics of the best rf model on the train set
rf_tune_results %>% 
  collect_metrics() %>%
  filter(.config == rf_best$.config)

#finalizing work flow
rf_wf <- Data_wf %>% add_model(rf_spec)
final_rf <- finalize_workflow(rf_wf, rf_best)

#performing the last fit
final_xgb_test_results <- last_fit(final_rf, Data_split, metrics = metrics)

# collecting the final predictions
collect_metrics(final_xgb_test_results)
```

```{r xgb_model}
set.seed(345)

#xgb tuneable model specifications
xgb_spec <- boost_tree(
  trees = 1000, 
  tree_depth = tune(),
  min_n = tune(), 
  loss_reduction = tune(),                     
  sample_size = tune(), mtry = tune(), ## randomness
  learn_rate = tune(), ## step size
  ) %>% 
  set_engine("xgboost") %>% 
  set_mode("classification")

#for tuning, a grid design that tries to maximize coverage of feature space
xgb_grid <- grid_latin_hypercube(
  tree_depth(),
  min_n(),
  loss_reduction(),
  sample_size = sample_prop(),
  finalize(mtry(), Data_train),
  learn_rate(),
  size = 40
)

#tuning procedure using k-folds cross validation
xgb_tune_results <- Data_wf %>%
  add_model(xgb_spec) %>%
  tune_grid(resamples = Data_folds,
  metrics = metrics,
  grid = xgb_grid,
  control = control_grid(save_pred = TRUE)
)

#visualizing tune results
xgb_tune_results %>%
  collect_metrics() %>%
  filter(.metric == "mn_log_loss") %>%
  dplyr::select(mean, mtry:sample_size) %>%
  pivot_longer(mtry:sample_size,
               values_to = "value",
               names_to = "parameter") %>%
  ggplot(aes(value, mean, color = parameter)) +
  geom_point(alpha = 0.8, show.legend = FALSE) +
  facet_wrap(~parameter, scales = "free_x") +
  labs(x = NULL, y = "mn_log_loss")

#selecting the best model
xgb_best <- xgb_tune_results %>% 
   select_best("mn_log_loss")

#collecting the predictions of the best model
#and mutating the column to identify the kind of model
xgb_train_best_pred <- xgb_tune_results %>% 
  collect_predictions(parameters = xgb_best) %>%
  mutate(model = "XGBoost")

#plotting the roc_curve for the best xgb model
xgb_train_best_pred %>% roc_curve(Sued, .pred_Yes) %>% autoplot()

#threshold assessment
#calculating the best threshold using the Fbeta measure
xgb_best_thresh_dat <- xgb_train_best_pred %>%
  filter(.config == xgb_best$.config) %>%
  pr_curve(Sued, .pred_Yes) %>%
  #F1-measure 
  mutate("F1 Measure" = ((1 + 1^2) * precision * recall) / ((1^2) * precision + recall)) %>%
  #F2-measure emphasizes minimizing false negatives
  mutate("F2 Measure" = ((1 + 2^2) * precision * recall) / ((2^2) * precision + recall)) %>%
  #F3-measure emphasizes minimizing false negatives even more
  mutate("F3 Measure" = ((1 + 2^3) * precision * recall) / ((2^3) * precision + recall)) %>%
  pivot_longer(!c(.threshold, recall, precision),
               names_to = "Fbeta_Measure", 
               values_to = "value")

#calculating ideal threshold using balanced accuracy
xgb_best_thresh_dat2 <- xgb_train_best_pred %>%
  filter(.config == xgb_best$.config) %>%
  roc_curve(Sued, .pred_Yes) %>%
  mutate("Bal_Acc" = (specificity + sensitivity)/2)

#saving the best threshold using balanced accuracy
xgb_best_thresh_BA <- xgb_best_thresh_dat2 %>% filter(Bal_Acc == max(Bal_Acc, na.rm = TRUE))
xgb_best_thresh_BA <- xgb_best_thresh_BA$.threshold

#saving the ideal threshold using F2 measure
xgb_best_thresh <- xgb_best_thresh_dat %>% filter(Fbeta_Measure == "F2 Measure") %>% filter(value == max(value, na.rm = TRUE))
xgb_best_thresh <- xgb_best_thresh$.threshold

#plot of thresholds
xgb_best_thresh_dat %>%
  ggplot(aes(x = .threshold, y = value, color = Fbeta_Measure)) +
  geom_path(show.legend = TRUE, alpha = 0.6, size = .75) +
  coord_equal() + 
  geom_vline(xintercept= max(xgb_best_thresh), 
             linetype="dashed", 
             color = "red",
             size = .5) +
  labs(title = "Optimal Classifier Threshold for XGBoost", y = "Value", x = "Threshold" )

#the confusion matrix of the best xgb model on the train set
xgb_tune_results %>% 
  collect_predictions() %>%
  filter(.config == xgb_best$.config) %>%
  #custom threshold
  mutate(.pred_class = ifelse(.pred_Yes > xgb_best_thresh, "Yes", "No")) %>%
  conf_mat(Sued, .pred_class)

#the metrics of the best xgb model on the train set
  xgb_tune_results %>% 
  collect_metrics() %>%
  filter(.config == xgb_best$.config)

#finalizing xgb work flow with the best hyper parameters
xgb_wf <- Data_wf %>% add_model(xgb_spec)
final_xgb <- finalize_workflow(xgb_wf, xgb_best)

#performing the last fit
final_xgb_test_results <- last_fit(final_xgb, Data_split, metrics = metrics)

#collecting the final predictions
collect_metrics(final_xgb_test_results)
```

```{r knn_model}
#might be necessary for random tie-breaking
set.seed(345)

#K-nearest neighbors spec
knn_spec <- 
  nearest_neighbor(neighbors = tune()) %>%
  set_engine("kknn") %>% 
  set_mode("classification")

#grid for tuning, only thing to tune is the # of neighbors
knn_grid <- tibble(neighbors = c(1:29, seq(from = 30, to = 100, by = 10), seq(from = 120, to = 200, by = 20)))

#tuning the model using k-folds cross validation
knn_tune_results <- Data_wf %>%
  add_model(knn_spec) %>% 
  tune_grid(Data_folds,
            grid = knn_grid,
            control = control_grid(save_pred = TRUE),
            metrics = metrics)

#data frame used to highlight the model with the lowest mn_log_loss
highlight_df <- knn_tune_results %>% 
  collect_metrics() %>% 
  filter(.metric == "mn_log_loss") %>%
  filter(mean == min(mean))

#plotting all models by mn_log_loss
knn_plot <- knn_tune_results %>% 
  collect_metrics() %>%
  filter(.metric == "mn_log_loss") %>%
  ggplot(aes(x = neighbors, y = mean)) + 
  #giving each model its distinct color
  geom_point(color = "#F8766D") + 
  ylab("Log Loss")+
  scale_x_log10(labels = scales::label_number()) +
  geom_line(color = "#F8766D") +
  geom_point(data=highlight_df, size = 3, color = "red2") +
  geom_vline(xintercept= highlight_df$neighbors, 
             linetype="dashed", 
             color = "red",
             size = .5)
knn_plot

#selecting the best model by highest AUC
knn_best <- select_best(knn_tune_results, metric = "mn_log_loss")
knn_best

#collecting the predictions of the best knn model
knn_train_best_pred <- knn_tune_results %>% 
  collect_predictions(parameters = knn_best) %>%
  mutate(model = "K-Nearest Neighbors")

#graphing the best model's ROC curve
knn_train_best_pred %>% roc_curve(Sued, .pred_Yes) %>% autoplot()

#threshold assessment for the confusion matrix
#calculating the ideal threshold using the Fbeta Measure
knn_best_thresh_dat <- knn_train_best_pred %>%
  filter(.config == knn_best$.config) %>%
  pr_curve(Sued, .pred_Yes) %>%
  #F1-measure 
  mutate("F1 Measure" = ((1 + 1^2) * precision * recall) / ((1^2) * precision + recall)) %>%
  #F2-measure emphasizes minimizing false negatives
  mutate("F2 Measure" = ((1 + 2^2) * precision * recall) / ((2^2) * precision + recall)) %>%
  #F3-measure emphasizes minimizing false negatives even more
  mutate("F3 Measure" = ((1 + 2^3) * precision * recall) / ((2^3) * precision + recall)) %>%
  pivot_longer(!c(.threshold, recall, precision),
               names_to = "Fbeta_Measure", 
               values_to = "value")

#calculating ideal threshold using balanced accuracy
knn_best_thresh_dat2 <- knn_train_best_pred %>%
  filter(.config == knn_best$.config) %>%
  roc_curve(Sued, .pred_Yes) %>%
  mutate("Bal_Acc" = (specificity + sensitivity)/2)

#saving the best threshold using balanced accuracy
knn_best_thresh_BA <- knn_best_thresh_dat2 %>% filter(Bal_Acc == max(Bal_Acc, na.rm = TRUE))
knn_best_thresh_BA <- knn_best_thresh_BA$.threshold

#saving the ideal threshold using F2 measure
knn_best_thresh <- knn_best_thresh_dat %>% filter(Fbeta_Measure == "F2 Measure") %>% filter(value == max(value, na.rm = TRUE))
knn_best_thresh <- knn_best_thresh$.threshold

#plot of thresholds
knn_best_thresh_dat %>%
  ggplot(aes(x = .threshold, y = value, color = Fbeta_Measure)) +
  geom_path(show.legend = TRUE, alpha = 0.6, size = .75) +
  coord_equal() + 
  geom_vline(xintercept= max(knn_best_thresh), 
             linetype="dashed", 
             color = "red",
             size = .5) +
  labs(title = "Optimal Classifier Threshold for K-Nearest Neighbors", y = "Value", x = "Threshold" )

#the confusion matrix of the best knn model on the train set (est using cv)
knn_tune_results %>% 
  collect_predictions() %>%
  filter(.config == knn_best$.config) %>%
  conf_mat(Sued, .pred_class)

#the metrics of the best knn model on the train set (est using cv)
  knn_tune_results %>% 
  collect_metrics() %>%
  filter(.config == knn_best$.config)
  
#finalizing knn work flow with the best hyper parameters
knn_wf <- Data_wf %>%
  add_model(knn_spec)
final_knn <- finalize_workflow(knn_wf, knn_best)

#performing the last fit
final_knn_test_results <- last_fit(final_knn, Data_split, metrics = metrics)

#collecting the final predictions
collect_metrics(final_knn_test_results)
```

#### Comparing the best model of each algorithm.

```{r Likelihood Model Training Dataset Graphs and Metrics}
#creating a data set of the predictions of the best models
Like_mdl_pred_train <- bind_rows(lr_train_best_pred,
                                 rf_train_best_pred,
                                 xgb_train_best_pred,
                                 knn_train_best_pred)

#comparing the roc of the best models
Like_mdl_pred_train %>%
  group_by(model) %>%
  roc_curve(Sued, .pred_Yes) %>%
  autoplot() + 
  geom_hline(yintercept= 1, 
             linetype="dashed", 
             color = "red",
             size = .5) +
  labs(title = "ROC Curve")

#comparing the pr curve of the best models
Like_mdl_pred_train %>%
  group_by(model) %>%
  pr_curve(Sued, .pred_Yes) %>%
  autoplot() + 
  labs(title = "PR Curve")

#det curve
Like_mdl_pred_train %>%
  group_by(model) %>%
  roc_curve(Sued, .pred_Yes) %>%
  ggplot(aes(x = 1 - specificity, y = 1 - sensitivity, color = model)) +
  geom_path() +
  labs(title = "Detection Error Tradeoff Curve", x = "False Positive Rate", y = "False Negative Rate") +
  scale_y_continuous(labels = scales::percent) +
  scale_x_continuous(labels = scales::percent) + 
  coord_cartesian(xlim=c(0,1), ylim=c(0,1),)

#But more important than a ranking metric like AUC is a probability metric like logloss

#create a data frame with all the best models' metrics (best by lowest mn_log_loss)
Like_mdl_met_train_ll <- bind_rows(lr_tune_results %>% collect_metrics(summarise = TRUE) %>% 
                                   dplyr::filter(.config == lr_tune_results %>% select_best("mn_log_loss") %>% pull()) %>% mutate(model = "Logistic Regression"),
                                 rf_tune_results %>% collect_metrics(summarise = TRUE) %>% 
                                   dplyr::filter(.config == rf_tune_results %>% select_best("mn_log_loss") %>% pull()) %>% mutate(model = "Random Forest"),
                                 xgb_tune_results %>% collect_metrics(summarise = TRUE) %>% 
                                   dplyr::filter(.config == xgb_tune_results %>% select_best("mn_log_loss") %>% pull()) %>% mutate(model = "XGBoost"),
                                 knn_tune_results %>% collect_metrics(summarise = TRUE) %>% 
                                   dplyr::filter(.config == knn_tune_results %>% select_best("mn_log_loss") %>% pull()) %>% mutate(model = "K-Nearest Neighbors")
                                 )

#comparing metrics of the best models
Like_mdl_met_train_ll <- 
  Like_mdl_met_train_ll %>% 
  dplyr::select(model, .metric, mean, std_err) %>% 
  pivot_wider(names_from = .metric, 
              values_from = c(mean, std_err)) 
Like_mdl_met_train_ll #%>% kable(booktabs = TRUE) %>% kable_styling(font_size = 10)

#comparing confusion matrices of the best models using the default .5 threshold
p1 <- lr_tune_results %>% 
  collect_predictions() %>%
  filter(.config == lr_best$.config) %>%
  conf_mat(Sued, .pred_class) %>%
  autoplot(cm, type = "heatmap") +
  labs(title = "Logistic Regression Confusion Matrix") +
  scale_fill_stepsn( colors = c("white", "#7CAE00"), breaks = c(35, 100, 300)) +
  theme(panel.border = element_rect(colour = "#7CAE00", fill=NA, size= 1))

p2 <- rf_tune_results %>% 
  collect_predictions() %>%
  filter(.config == rf_best$.config) %>%
  conf_mat(Sued, .pred_class) %>%
  autoplot(cm, type = "heatmap") +
  labs(title = "Random Forest Confusion Matrix") +
  scale_fill_stepsn( colors = c("white", "#00BFC4"), breaks = c(35, 100, 300)) +
  theme(panel.border = element_rect(colour = "#00BFC4", fill=NA, size= 1))

p3 <- xgb_tune_results %>% 
  collect_predictions() %>%
  filter(.config == xgb_best$.config) %>%
  conf_mat(Sued, .pred_class) %>%
  autoplot(cm, type = "heatmap") +
  labs(title = "XGBoost Confusion Matrix") +
  scale_fill_stepsn( colors = c("white", "#C77CFF"), breaks = c(35, 100, 300)) +
  theme(panel.border = element_rect(colour = "#C77CFF", fill=NA, size= 1))

p4 <- knn_tune_results %>%
  collect_predictions() %>%
  filter(.config == knn_best$.config) %>%
  conf_mat(Sued, .pred_class) %>%
  autoplot(cm, type = "heatmap") +
  labs(title = "K-NN Confusion Matrix") +
  scale_fill_stepsn( colors = c("white", "#F8766D"), breaks = c(35, 100, 300)) +
  theme(panel.border = element_rect(colour = "#F8766D", fill=NA, size= 1))

#comparing confusion matrices of the best models using the best threshold by
#cross validated F2 measure
p5 <- lr_tune_results %>% 
  collect_predictions() %>%
  filter(.config == lr_best$.config) %>%
  #custom threshold
  mutate(.pred_class = ifelse(.pred_Yes > lr_best_thresh, "Yes", "No")) %>%
  conf_mat(Sued, .pred_class) %>%
  autoplot(cm, type = "heatmap") +
  labs(title = "Logistic Regression Confusion Matrix") +
  scale_fill_stepsn( colors = c("white", "#7CAE00"), breaks = c(35, 100, 300)) +
  theme(panel.border = element_rect(colour = "#7CAE00", fill=NA, size= 1))

p6 <- rf_tune_results %>% 
  collect_predictions() %>%
  filter(.config == rf_best$.config) %>%
  #custom threshold
  mutate(.pred_class = ifelse(.pred_Yes > rf_best_thresh, "Yes", "No")) %>%
  conf_mat(Sued, .pred_class) %>%
  autoplot(cm, type = "heatmap") +
  labs(title = "Random Forest Confusion Matrix") +
  scale_fill_stepsn( colors = c("white", "#00BFC4"), breaks = c(35, 100, 300)) +
  theme(panel.border = element_rect(colour = "#00BFC4", fill=NA, size= 1))

p7 <- xgb_tune_results %>% 
  collect_predictions() %>%
  filter(.config == xgb_best$.config) %>%
  #custom threshold
  mutate(.pred_class = ifelse(.pred_Yes > xgb_best_thresh, "Yes", "No")) %>%
  conf_mat(Sued, .pred_class) %>%
  autoplot(cm, type = "heatmap") +
  labs(title = "XGBoost Confusion Matrix") +
  scale_fill_stepsn( colors = c("white", "#C77CFF"), breaks = c(35, 100, 300)) +
  theme(panel.border = element_rect(colour = "#C77CFF", fill=NA, size= 1))

p8 <- knn_tune_results %>%
  collect_predictions() %>%
  filter(.config == knn_best$.config) %>%
  #custom threshold
  mutate(.pred_class = ifelse(.pred_Yes > knn_best_thresh, "Yes", "No")) %>%
  conf_mat(Sued, .pred_class) %>%
  autoplot(cm, type = "heatmap") +
  labs(title = "K-NN Confusion Matrix") +
  scale_fill_stepsn( colors = c("white", "#F8766D"), breaks = c(35, 100, 300)) +
  theme(panel.border = element_rect(colour = "#F8766D", fill=NA, size= 1))

#comparing confusion matrices of the best models using the best threshold by
#cross validated Balanced accuracy
p9 <- lr_tune_results %>% 
  collect_predictions() %>%
  filter(.config == lr_best$.config) %>%
  #custom threshold
  mutate(.pred_class = ifelse(.pred_Yes > lr_best_thresh_BA, "Yes", "No")) %>%
  conf_mat(Sued, .pred_class) %>%
  autoplot(cm, type = "heatmap") +
  labs(title = "Logistic Regression Confusion Matrix") +
  scale_fill_stepsn( colors = c("white", "#7CAE00"), breaks = c(35, 100, 300)) +
  theme(panel.border = element_rect(colour = "#7CAE00", fill=NA, size= 1))

p10 <- rf_tune_results %>% 
  collect_predictions() %>%
  filter(.config == rf_best$.config) %>%
  #custom threshold
  mutate(.pred_class = ifelse(.pred_Yes > rf_best_thresh_BA, "Yes", "No")) %>%
  conf_mat(Sued, .pred_class) %>%
  autoplot(cm, type = "heatmap") +
  labs(title = "Random Forest Confusion Matrix") +
  scale_fill_stepsn( colors = c("white", "#00BFC4"), breaks = c(35, 100, 300)) +
  theme(panel.border = element_rect(colour = "#00BFC4", fill=NA, size= 1))

p11 <- xgb_tune_results %>% 
  collect_predictions() %>%
  filter(.config == xgb_best$.config) %>%
  #custom threshold
  mutate(.pred_class = ifelse(.pred_Yes > xgb_best_thresh_BA, "Yes", "No")) %>%
  conf_mat(Sued, .pred_class) %>%
  autoplot(cm, type = "heatmap") +
  labs(title = "XGBoost Confusion Matrix") +
  scale_fill_stepsn( colors = c("white", "#C77CFF"), breaks = c(35, 100, 300)) +
  theme(panel.border = element_rect(colour = "#C77CFF", fill=NA, size= 1))

p12 <- knn_tune_results %>%
  collect_predictions() %>%
  filter(.config == knn_best$.config) %>%
  #custom threshold
  mutate(.pred_class = ifelse(.pred_Yes > knn_best_thresh_BA, "Yes", "No")) %>%
  conf_mat(Sued, .pred_class) %>%
  autoplot(cm, type = "heatmap") +
  labs(title = "K-NN Confusion Matrix") +
  scale_fill_stepsn( colors = c("white", "#F8766D"), breaks = c(35, 100, 300)) +
  theme(panel.border = element_rect(colour = "#F8766D", fill=NA, size= 1))

#plotting the confusion matrices with different cross-validated thresholds
grid.arrange(p1, p2, p3, p4, nrow = 2)
grid.arrange(p5, p6, p7, p8, nrow = 2)
grid.arrange(p9, p10, p11, p12, nrow = 2)

#calculating the log loss of the reference model
ll_baseline <- rf_tune_results %>% 
  collect_predictions() %>%
  filter(.config == rf_best$.config) %>%
  mutate(baseline_pred = sum(Data_train$Sued == "Yes")/sum(Data_train$Sued == "No")) %>%
  mn_log_loss(estimate = baseline_pred, truth = Sued) %>%
  pull(.estimate)

#all models beat the baseline
Like_mdl_met_train_ll %>% 
  arrange(mean_mn_log_loss, decreasing = TRUE) %>%
  bind_rows(data.frame(model= "Non-Informative Baseline", mean_mn_log_loss = ll_baseline)) %>%
  mutate(model = fct_reorder(model, mean_mn_log_loss, .desc = TRUE)) %>%
  ggplot(aes(model, mean_mn_log_loss, fill=model)) +
  geom_col(show.legend = FALSE) +
  scale_fill_manual(values = c("K-Nearest Neighbors" = "#F8766D",
                               "Logistic Regression"="#7CAE00",
                               "Random Forest"="#00BFC4",
                               "XGBoost"="#C77CFF",
                               "Non-Informative Baseline" = "gray")) +
  coord_flip() +
  geom_text(size = 3,
               aes(label = round(mean_mn_log_loss, 3), y = mean_mn_log_loss + .01),
               vjust = 1) +
  labs(title = "Best Model by Cross-Validated Log Loss", x = "Model", y = "Log Loss")
```

#### Using the test set to assess generalization performance of the best model.

```{r Best Likelihood Model Test Dataset Graphs and Metrics}
#collecting predictions on the test data of the final model
#and mutating a column to identify model type
#creating a dataframe with all the best models metrics
Like_mdl_met_test <- bind_rows(final_lr_test_results %>% collect_metrics(summarise = TRUE) %>% mutate(model = "Logistic Regression"),
                               final_xgb_test_results %>% collect_metrics(summarise = TRUE) %>% mutate(model = "Random Forest"),
                               final_xgb_test_results %>% collect_metrics(summarise = TRUE) %>% mutate(model = "XGBoost"),
                               final_knn_test_results %>% collect_metrics(summarise = TRUE) %>% mutate(model = "K-Nearest Neighbors")
                               )


#looking at the metrics of the final model
Like_mdl_met_test <- 
  Like_mdl_met_test %>%
  dplyr::select(model, .metric, .estimate) %>% 
  pivot_wider(names_from = .metric, 
              values_from = .estimate)
Like_mdl_met_test

#plotting the CM of the test set pred of the final model
final_xgb_test_results %>%
  collect_predictions() %>%
  #custom threshold
  mutate(.pred_class = ifelse(.pred_Yes > xgb_best_thresh_BA, "Yes", "No")) %>%
  conf_mat(Sued, .pred_class) %>%
  autoplot(cm, type = "heatmap") +
  labs(title = "XGB Confusion Matrix on Test with Optimal Balanced Accuracy as Threshold") +
  scale_fill_stepsn( colors = c("white", "#C77CFF"), breaks = c(15, 40, 100)) +
  theme(panel.border = element_rect(colour = "#C77CFF", fill=NA, size= 1))

#fitting the final model and making an uncalibrated prediction
final_xgb_model <- fit(final_xgb, Data_train)
predict(final_xgb_model, Mohawk_Obs, type = "prob") %>% mutate(model = "XGBoost")

#looking at feature importance
final_xgb_model %>% 
  extract_fit_parsnip() %>%
  vip(num_features = 6)
```

#### Calibrating the best model using k-folds cross validation and assessing the calibrated model's performance on the test set.

```{r Final Model Calibration Assessment and Adjustment}
#Okay it looks like xgb is the best model
#but since we are looking at probabilities (and not labels or rankings)
#we need to assess model calibration, adjust model calibration if necessary

#assessing calibration using k-folds cross validation
xgb_cv <- final_xgb %>%
  fit_resamples(resamples = Data_cal_folds, control = control_resamples(save_pred = TRUE, save_workflow = TRUE))
xgb_cv <- xgb_cv %>% collect_predictions()

#knn_mccv <- final_knn %>% 
#  fit_resamples(resamples = Data_cal_folds, control = control_resamples(save_pred = TRUE, save_workflow = TRUE))
#knn_mccv <- knn_mccv %>% collect_predictions()

#looking at the probability distribution of the rf model 
#which was looking like the best

ggplot(xgb_cv, aes(x=.pred_Yes, fill = Sued, alpha = .5)) + 
  geom_density()
summary(xgb_cv$.pred_Yes)

#xgb_cv$.pred_Yes <- classadjust(xgb_cv$.pred_Yes, .5, sum(Data$Sued == "Yes")/sum(Data$Sued == "No"))

#one hot encoding Sued for plotting the calibration curve
#lr_mccv$Sued_1hot <- ifelse(lr_mccv$Sued == "Yes", 1, 0)
xgb_cv$Sued_1hot <- ifelse(xgb_cv$Sued == "Yes", 1, 0)
#xgb_mccv$Sued_1hot <- ifelse(xgb_mccv$Sued == "Yes", 1, 0)
#knn_mccv$Sued_1hot <- ifelse(knn_mccv$Sued == "Yes", 1, 0)

#summary(verify(obs = lr_mccv$Sued_1hot, pred = lr_mccv$.pred_Yes))
summary(verify(obs = xgb_cv$Sued_1hot, pred = xgb_cv$.pred_Yes))
#summary(verify(obs = xgb_mccv$Sued_1hot, pred = xgb_mccv$.pred_Yes))
#summary(verify(obs = knn_mccv$Sued_1hot, pred = knn_mccv$.pred_Yes))


#performing iso reg calibration
xgb_cv$iso_Yes <- probability.calibration(xgb_cv$Sued_1hot, xgb_cv$.pred_Yes)

#performing beta calibration
 beta_cal <- beta_calibration(xgb_cv$.pred_Yes, xgb_cv$Sued_1hot, "abm")
 xgb_cv$beta_Yes <- beta_predict(xgb_cv$.pred_Yes, beta_cal ) 

# performing platt scaling on the data set
plattdf <-data.frame(xgb_cv$.pred_Yes, xgb_cv$Sued_1hot)
colnames(plattdf)<-c("x","y")
# training a logistic regression model on the cross validation dataset
rf_platt_model<-glm(y~x,data = plattdf,family = binomial)
#predicting on the cross validation after platt scaling
platt_Yes <- predict(rf_platt_model, plattdf[-2], type = "response")
xgb_cv$platt_Yes <- platt_Yes

#calculating log loss baseline (non informative value)
xgb_cv$bad <- sum(Data_train$Sued == "Yes")/sum(Data_train$Sued == "No")
lldf1 <- mn_log_loss(data = xgb_cv, estimate = bad, truth = Sued) %>% mutate(calibration = "reference model")
#calculating log loss before calibration
lldf2 <- mn_log_loss(data = xgb_cv, estimate = .pred_Yes, truth = Sued) %>% mutate(calibration = "none")
#calculating log loss after platt scaling calibration
lldf3 <- mn_log_loss(data = xgb_cv, estimate = platt_Yes, truth = Sued) %>% mutate(calibration = "platt scaling")
#calculating log loss after iso reg calibration
lldf4 <- mn_log_loss(data = xgb_cv, estimate = iso_Yes, truth = Sued) %>% mutate(calibration = "isotonic regression")
#calculating log loss after beta calibration
lldf5 <- mn_log_loss(data = xgb_cv, estimate = beta_Yes, truth = Sued) %>% mutate(calibration = "beta calibration")
#binding and displaying comparison
ll_cal_compare <- rbind(lldf1, lldf2, lldf3, lldf5)
ll_cal_compare

#tested Brier skill score since it is like r squared
#but it is not the best for infrequent events
#but is better for interpretability
summary(verify(obs = xgb_cv$Sued_1hot, pred = xgb_cv$.pred_Yes))
summary(verify(obs = xgb_cv$Sued_1hot, pred = xgb_cv$platt_Yes))
summary(verify(obs = xgb_cv$Sued_1hot, pred = xgb_cv$iso_Yes))
summary(verify(obs = xgb_cv$Sued_1hot, pred = xgb_cv$beta_Yes))

############################ TEST ##############################################
    
#looking at the calibration of the model on the held out set
#as the final depiction of model performance
xgb_test_res <- final_xgb_test_results %>% collect_predictions()
xgb_test_res$Sued_1hot <- ifelse(xgb_test_res$Sued == "Yes", 1, 0)

#performing platt scaling on test
plattdf_test <-data.frame(xgb_test_res$.pred_Yes, xgb_test_res$Sued_1hot)
colnames(plattdf_test)<-c("x","y")
xgb_test_res$platt_Yes <- predict(rf_platt_model, plattdf_test[-2], type = "response")

#performing beta cal on test
xgb_test_res$beta_Yes <- beta_predict(xgb_test_res$.pred_Yes, beta_cal) 

#score assessment
rbind(mn_log_loss(data = xgb_test_res, estimate = .pred_Yes, truth = Sued) %>% mutate(calibration = "none"),
mn_log_loss(data = xgb_test_res, estimate = platt_Yes, truth = Sued) %>% mutate(calibration = "platt"),
mn_log_loss(data = xgb_test_res, estimate = beta_Yes, truth = Sued) %>% mutate(calibration = "beta"))
summary(verify(obs = xgb_test_res$Sued_1hot, pred = xgb_test_res$.pred_Yes))
summary(verify(obs = xgb_test_res$Sued_1hot, pred = xgb_test_res$platt_Yes))
summary(verify(obs = xgb_test_res$Sued_1hot, pred = xgb_test_res$beta_Yes))
```

#### Graphing calibration curves on the calibration set (cross validation) and test set.

```{r Calibration Curve Function}
#modifying code from remotes::install_github('ML4LHS/runway')
#to make calibration plots

#function that plots calibration curves
#pred_data is the data set that contains predictions
#pred_var is the variable in pred_data that contains probability predictions
#turth_var is the variable in pred_data that contains the one hot encoded true class
plot_cal_curve <- function(pred_data, pred_var, truth_var){
  
  #creating the bins and standard errors
  cal_plot <- dplyr::mutate(pred_data, bin = dplyr::ntile(pred_var, 10)) %>% 
  dplyr::group_by(bin) %>% 
  dplyr::mutate(n = dplyr::n(), 
                bin_pred = mean(pred_var, na.rm = TRUE), 
                bin_prob = mean(truth_var, na.rm = TRUE), 
                se = sqrt((bin_prob * (1 - bin_prob))/n), 
                ul = bin_prob + 1.96 * se, 
                ll = bin_prob - 1.96 * se) %>% 
  dplyr::mutate_at(dplyr::vars(ul, ll), . %>% scales::oob_squish(range = c(0, 1))) %>% 
  dplyr::ungroup()

  #plotting the calibration curve
  cal_plot %>% 
  ggplot2::ggplot() + 
  ggplot2::scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, by = .1)) + 
  ggplot2::scale_x_continuous(limits = c(0, 1), breaks = seq(0, 1, by = 0.1)) + 
  ggplot2::geom_abline(linetype = "dashed") + 
  ggplot2::geom_point(ggplot2::aes(x = bin_pred, y = bin_prob, ymin = ll, ymax = ul), 
                      size = 2, 
                      color = "black") + 
  ggplot2::geom_errorbar(ggplot2::aes(x = bin_pred, y = bin_prob, ymin = ll, ymax = ul), 
                         size = 0.5, 
                         color = "black", width = 0.02) + 
  ggplot2::geom_point(ggplot2::aes(x = bin_pred, y = bin_prob, ymin = ll, ymax = ul), 
                      size = 2, 
                      color = "black") + 
  ggplot2::geom_errorbar(ggplot2::aes(x = bin_pred, y = bin_prob, ymin = ll, ymax = ul), 
                         size = 0.5, 
                         color = "black", 
                         width = 0.02) + 
  ggplot2::geom_smooth(ggplot2::aes(x = pred_var, y = truth_var, color = "#00BFC4"), 
                       se = TRUE, 
                       method = "loess") + 
  labs(title = "XGBoost Calibration curve", 
       x = "Predicted probability",
       y = "Empirical probability") +
  ggplot2::theme_minimal() + 
  ggplot2::theme(aspect.ratio = 1, legend.position="none") +
  #modify if you want to zoom in on part of the plot
  coord_cartesian(xlim=c(0,1), ylim=c(0,1),)
}

plot_cal_curve(xgb_cv, .pred_Yes, Sued_1hot)
```

```{r Likelihood Calibration Curves Validation Set}
#modifying code from remotes::install_github('ML4LHS/runway')
#to make calibration plots
#creating the bins
rf_cal_plot <- dplyr::mutate(xgb_cv, bin = dplyr::ntile(.pred_Yes, 10)) %>% 
  dplyr::group_by(bin) %>% 
  dplyr::mutate(n = dplyr::n(), 
                bin_pred = mean(.pred_Yes, na.rm = TRUE), 
                bin_prob = mean(Sued_1hot, na.rm = TRUE), 
                se = sqrt((bin_prob * (1 - bin_prob))/n), ul = bin_prob + 1.96 * se, ll = bin_prob - 1.96 * se) %>% 
  dplyr::mutate_at(dplyr::vars(ul, ll), . %>% scales::oob_squish(range = c(0, 1))) %>% 
  dplyr::ungroup()
#plotting
  rf_cal_plot %>% 
  ggplot2::ggplot() + 
  ggplot2::scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, by = .1)) + 
  ggplot2::scale_x_continuous(limits = c(0, 1), breaks = seq(0, 1, by = 0.1)) + 
  ggplot2::geom_abline(linetype = "dashed") + 
  ggplot2::geom_point(ggplot2::aes(x = bin_pred, y = bin_prob, ymin = ll, ymax = ul), 
                      size = 2, 
                      color = "black") + 
  ggplot2::geom_errorbar(ggplot2::aes(x = bin_pred, y = bin_prob, ymin = ll, ymax = ul), 
                         size = 0.5, 
                         color = "black", width = 0.02) + 
  ggplot2::geom_point(ggplot2::aes(x = bin_pred, y = bin_prob, ymin = ll, ymax = ul), 
                      size = 2, 
                      color = "black") + 
  ggplot2::geom_errorbar(ggplot2::aes(x = bin_pred, y = bin_prob, ymin = ll, ymax = ul), 
                         size = 0.5, 
                         color = "black", 
                         width = 0.02) + 
  ggplot2::geom_smooth(ggplot2::aes(x = .pred_Yes, y = Sued_1hot, color = "#00BFC4"), se = TRUE, method = "loess") + 
  labs(title = "XGBoost Calibration curve", 
     x = "Predicted probability",
     y = "Empirical probability") +
  ggplot2::theme_minimal() + 
  ggplot2::theme(aspect.ratio = 1, legend.position="none") +
  coord_cartesian(xlim=c(0,1), ylim=c(0,1),)
  
  #####################  PLATT
  
  rf_cal_plot <- dplyr::mutate(xgb_cv, bin = dplyr::ntile(platt_Yes, 10)) %>% 
  dplyr::group_by(bin) %>% 
  dplyr::mutate(n = dplyr::n(), 
                bin_pred = mean(platt_Yes, na.rm = TRUE), 
                bin_prob = mean(Sued_1hot, na.rm = TRUE), 
                se = sqrt((bin_prob * (1 - bin_prob))/n), ul = bin_prob + 1.96 * se, ll = bin_prob - 1.96 * se) %>% 
  dplyr::mutate_at(dplyr::vars(ul, ll), . %>% scales::oob_squish(range = c(0, 1))) %>% 
  dplyr::ungroup()
#plotting
  rf_cal_plot %>% 
  ggplot2::ggplot() + 
  ggplot2::scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, by = .1)) + 
  ggplot2::scale_x_continuous(limits = c(0, 1), breaks = seq(0, 1, by = 0.1)) + 
  ggplot2::geom_abline(linetype = "dashed") + 
  ggplot2::geom_point(ggplot2::aes(x = bin_pred, y = bin_prob, ymin = ll, ymax = ul), 
                      size = 2, 
                      color = "black") + 
  ggplot2::geom_errorbar(ggplot2::aes(x = bin_pred, y = bin_prob, ymin = ll, ymax = ul), 
                         size = 0.5, 
                         color = "black", width = 0.02) + 
  ggplot2::geom_point(ggplot2::aes(x = bin_pred, y = bin_prob, ymin = ll, ymax = ul), 
                      size = 2, 
                      color = "black") + 
  ggplot2::geom_errorbar(ggplot2::aes(x = bin_pred, y = bin_prob, ymin = ll, ymax = ul), 
                         size = 0.5, 
                         color = "black", 
                         width = 0.02) + 
  ggplot2::geom_smooth(ggplot2::aes(x = platt_Yes, y = Sued_1hot, color = "#00BFC4"), se = TRUE, method = "loess") + 
  labs(title = "XGBoost Calibration curve (Platt Scaling)", 
     x = "Predicted probability",
     y = "Empirical probability") +
  ggplot2::theme_minimal() + 
  ggplot2::theme(aspect.ratio = 1, legend.position="none") +
  coord_cartesian(xlim=c(0,1), ylim=c(0,1),)
  
  ###############################  ISOREG
  
  rf_cal_plot <- dplyr::mutate(xgb_cv, bin = dplyr::ntile(iso_Yes, 10)) %>% 
  dplyr::group_by(bin) %>% 
  dplyr::mutate(n = dplyr::n(), 
                bin_pred = mean(iso_Yes, na.rm = TRUE), 
                bin_prob = mean(Sued_1hot, na.rm = TRUE), 
                se = sqrt((bin_prob * (1 - bin_prob))/n), 
                ul = bin_prob + 1.96 * se, 
                ll = bin_prob - 1.96 * se) %>% 
  dplyr::mutate_at(dplyr::vars(ul, ll), . %>% scales::oob_squish(range = c(0, 1))) %>% 
  dplyr::ungroup()
#plotting
  rf_cal_plot %>% 
  ggplot2::ggplot() + 
  ggplot2::scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, by = .1)) + 
  ggplot2::scale_x_continuous(limits = c(0, 1), breaks = seq(0, 1, by = .1)) + 
  ggplot2::geom_abline(linetype = "dashed") + 
  ggplot2::geom_point(ggplot2::aes(x = bin_pred, y = bin_prob, ymin = ll, ymax = ul), 
                      size = 2, 
                      color = "black") + 
  ggplot2::geom_errorbar(ggplot2::aes(x = bin_pred, y = bin_prob, ymin = ll, ymax = ul), 
                         size = 0.5, 
                         color = "black", width = 0.02) + 
  ggplot2::geom_point(ggplot2::aes(x = bin_pred, y = bin_prob, ymin = ll, ymax = ul), 
                      size = 2, 
                      color = "black") + 
  ggplot2::geom_errorbar(ggplot2::aes(x = bin_pred, y = bin_prob, ymin = ll, ymax = ul), 
                         size = 0.5, 
                         color = "black", 
                         width = 0.02) + 
  ggplot2::geom_smooth(ggplot2::aes(x = iso_Yes, y = Sued_1hot, color = "#00BFC4"), se = TRUE, method = "loess") + 
  labs(title = "XGBoost Calibration curve (Isotonic Regression)", 
     x = "Predicted probability",
     y = "Empirical probability") +
  ggplot2::theme_minimal() + 
  ggplot2::theme(aspect.ratio = 1, legend.position="none") +
  coord_cartesian(xlim=c(0,1), ylim=c(0,1),)
  
###############################  BETA
  
  rf_cal_plot <- dplyr::mutate(xgb_cv, bin = dplyr::ntile(beta_Yes, 10)) %>% 
  dplyr::group_by(bin) %>% 
  dplyr::mutate(n = dplyr::n(), 
                bin_pred = mean(beta_Yes, na.rm = TRUE), 
                bin_prob = mean(Sued_1hot, na.rm = TRUE), 
                se = sqrt((bin_prob * (1 - bin_prob))/n), ul = bin_prob + 1.96 * se, ll = bin_prob - 1.96 * se) %>% 
  dplyr::mutate_at(dplyr::vars(ul, ll), . %>% scales::oob_squish(range = c(0, 1))) %>% 
  dplyr::ungroup()
#plotting
  rf_cal_plot %>% 
  ggplot2::ggplot() + 
  ggplot2::scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, by = .1)) + 
  ggplot2::scale_x_continuous(limits = c(0, 1), breaks = seq(0, 1, by = .1)) + 
  ggplot2::geom_abline(linetype = "dashed") + 
  ggplot2::geom_point(ggplot2::aes(x = bin_pred, y = bin_prob, ymin = ll, ymax = ul), 
                      size = 2, 
                      color = "black") + 
  ggplot2::geom_errorbar(ggplot2::aes(x = bin_pred, y = bin_prob, ymin = ll, ymax = ul), 
                         size = 0.5, 
                         color = "black", width = 0.02) + 
  ggplot2::geom_point(ggplot2::aes(x = bin_pred, y = bin_prob, ymin = ll, ymax = ul), 
                      size = 2, 
                      color = "black") + 
  ggplot2::geom_errorbar(ggplot2::aes(x = bin_pred, y = bin_prob, ymin = ll, ymax = ul), 
                         size = 0.5, 
                         color = "black", 
                         width = 0.02) + 
  ggplot2::geom_smooth(ggplot2::aes(x = beta_Yes, y = Sued_1hot, color = "#00BFC4"), se = TRUE, method = "loess") + 
  labs(title = "XGBoost Calibration curve (Beta Calibration)", 
     x = "Predicted probability",
     y = "Empirical probability") +
  ggplot2::theme_minimal() + 
  ggplot2::theme(aspect.ratio = 1, legend.position="none") +
  coord_cartesian(xlim=c(0,1), ylim=c(0,1),)
```

```{r Likelihood Calibration Curves Test Set}
#calibration plot on test w/out scaling
rf_cal_plot <- dplyr::mutate(xgb_test_res, bin = dplyr::ntile(.pred_Yes, 10)) %>% 
  dplyr::group_by(bin) %>% 
  dplyr::mutate(n = dplyr::n(), 
                bin_pred = mean(.pred_Yes, na.rm = TRUE), 
                bin_prob = mean(Sued_1hot, na.rm = TRUE), 
                se = sqrt((bin_prob * (1 - bin_prob))/n), ul = bin_prob + 1.96 * se, ll = bin_prob - 1.96 * se) %>% 
  dplyr::mutate_at(dplyr::vars(ul, ll), . %>% scales::oob_squish(range = c(0, 1))) %>% 
  dplyr::ungroup()
#plotting
  rf_cal_plot %>% 
  ggplot2::ggplot() + 
  ggplot2::scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, by = .1)) + 
  ggplot2::scale_x_continuous(limits = c(0, 1), breaks = seq(0, 1, by = 0.1)) + 
  ggplot2::geom_abline(linetype = "dashed") + 
  ggplot2::geom_point(ggplot2::aes(x = bin_pred, y = bin_prob, ymin = ll, ymax = ul), 
                      size = 2, 
                      color = "black") + 
  ggplot2::geom_errorbar(ggplot2::aes(x = bin_pred, y = bin_prob, ymin = ll, ymax = ul), 
                         size = 0.5, 
                         color = "black", width = 0.02) + 
  ggplot2::geom_point(ggplot2::aes(x = bin_pred, y = bin_prob, ymin = ll, ymax = ul), 
                      size = 2, 
                      color = "black") + 
  ggplot2::geom_errorbar(ggplot2::aes(x = bin_pred, y = bin_prob, ymin = ll, ymax = ul), 
                         size = 0.5, 
                         color = "black", 
                         width = 0.02) + 
  ggplot2::geom_smooth(ggplot2::aes(x = .pred_Yes, y = Sued_1hot, color = "#00BFC4"), se = TRUE, method = "loess") + 
  labs(title = "XGBoost Calibration Curve on Test", 
     x = "Predicted probability",
     y = "Empirical probability") +
  ggplot2::theme_minimal() + 
  ggplot2::theme(aspect.ratio = 1, legend.position="none") +
  coord_cartesian(xlim=c(0,1), ylim=c(0,1),)
  
#calibration plot on test w/ platt scaling
rf_cal_plot <- dplyr::mutate(xgb_test_res, bin = dplyr::ntile(platt_Yes, 10)) %>% 
  dplyr::group_by(bin) %>% 
  dplyr::mutate(n = dplyr::n(), 
                bin_pred = mean(platt_Yes, na.rm = TRUE), 
                bin_prob = mean(Sued_1hot, na.rm = TRUE), 
                se = sqrt((bin_prob * (1 - bin_prob))/n), ul = bin_prob + 1.96 * se, ll = bin_prob - 1.96 * se) %>% 
  dplyr::mutate_at(dplyr::vars(ul, ll), . %>% scales::oob_squish(range = c(0, 1))) %>% 
  dplyr::ungroup()
#plotting
  rf_cal_plot %>% 
  ggplot2::ggplot() + 
  ggplot2::scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, by = .1)) + 
  ggplot2::scale_x_continuous(limits = c(0, 1), breaks = seq(0, 1, by = 0.1)) + 
  ggplot2::geom_abline(linetype = "dashed") + 
  ggplot2::geom_point(ggplot2::aes(x = bin_pred, y = bin_prob, ymin = ll, ymax = ul), 
                      size = 2, 
                      color = "black") + 
  ggplot2::geom_errorbar(ggplot2::aes(x = bin_pred, y = bin_prob, ymin = ll, ymax = ul), 
                         size = 0.5, 
                         color = "black", width = 0.02) + 
  ggplot2::geom_point(ggplot2::aes(x = bin_pred, y = bin_prob, ymin = ll, ymax = ul), 
                      size = 2, 
                      color = "black") + 
  ggplot2::geom_errorbar(ggplot2::aes(x = bin_pred, y = bin_prob, ymin = ll, ymax = ul), 
                         size = 0.5, 
                         color = "black", 
                         width = 0.02) + 
  ggplot2::geom_smooth(ggplot2::aes(x = platt_Yes, y = Sued_1hot, color = "#00BFC4"), se = TRUE, method = "loess") + 
  labs(title = "XGBoost Calibration Curve on Test (Platt)", 
     x = "Predicted probability",
     y = "Empirical probability") +
  ggplot2::theme_minimal() + 
  ggplot2::theme(aspect.ratio = 1, legend.position="none") +
  coord_cartesian(xlim=c(0,1), ylim=c(0,1),)
  
#calibration plot on test w/ beta cal
rf_cal_plot <- dplyr::mutate(xgb_test_res, bin = dplyr::ntile(beta_Yes, 10)) %>% 
  dplyr::group_by(bin) %>% 
  dplyr::mutate(n = dplyr::n(), 
                bin_pred = mean(beta_Yes, na.rm = TRUE), 
                bin_prob = mean(Sued_1hot, na.rm = TRUE), 
                se = sqrt((bin_prob * (1 - bin_prob))/n), ul = bin_prob + 1.96 * se, ll = bin_prob - 1.96 * se) %>% 
  dplyr::mutate_at(dplyr::vars(ul, ll), . %>% scales::oob_squish(range = c(0, 1))) %>% 
  dplyr::ungroup()
#plotting
  rf_cal_plot %>% 
  ggplot2::ggplot() + 
  ggplot2::scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, by = .1)) + 
  ggplot2::scale_x_continuous(limits = c(0, 1), breaks = seq(0, 1, by = 0.1)) + 
  ggplot2::geom_abline(linetype = "dashed") + 
  ggplot2::geom_point(ggplot2::aes(x = bin_pred, y = bin_prob, ymin = ll, ymax = ul), 
                      size = 2, 
                      color = "black") + 
  ggplot2::geom_errorbar(ggplot2::aes(x = bin_pred, y = bin_prob, ymin = ll, ymax = ul), 
                         size = 0.5, 
                         color = "black", width = 0.02) + 
  ggplot2::geom_point(ggplot2::aes(x = bin_pred, y = bin_prob, ymin = ll, ymax = ul), 
                      size = 2, 
                      color = "black") + 
  ggplot2::geom_errorbar(ggplot2::aes(x = bin_pred, y = bin_prob, ymin = ll, ymax = ul), 
                         size = 0.5, 
                         color = "black", 
                         width = 0.02) + 
  ggplot2::geom_smooth(ggplot2::aes(x = beta_Yes, y = Sued_1hot, color = "#00BFC4"), se = TRUE, method = "loess") + 
  labs(title = "XGBoost Calibration Curve on Test (Beta)", 
     x = "Predicted probability",
     y = "Empirical probability") +
  ggplot2::theme_minimal() + 
  ggplot2::theme(aspect.ratio = 1, legend.position="none") +
  coord_cartesian(xlim=c(0,1), ylim=c(0,1),)
```

#### Using the calibrated best model to make the final prediction of SCA litigation likelihood for Mohawk Industries Inc.

```{r Best Likelihood Prediction}
#final prediction with the best performing beta calibration ensemble
likelihood_pred <- predict(final_xgb_model, Mohawk_Obs, type = "prob") %>% mutate(model = "XGBoost")
beta_predict(likelihood_pred$.pred_Yes, beta_cal) 
```

## Severity Modeling

#### Predicting the cost of SCA litigation for Mohawk Industries Inc.

The creation of a high quality severity model necessitated optimizing model hyperparameters and comparing multiple algorithms.In order to optimize hyperparameters and compare algorithms a nested cross validation procedure was implemented since the data set was relatively small (n\<100), thus utilizing the data budget optimally was essential. Nested cross-validation entails "nesting" a cross-validation loop, referred to as the inner loop, for hyperparameter optimization within another cross-validation loop, referred to as the outer loop, for estimating generalization accuracy. This procedure for model comparison optimally navigates the training and test data split trade-off that is especially detrimental to small data sets and circumvents the optimization bias introduced by using the same k-fold cross validation procedure to both tune model hyperparameters and evaluate models. In this case the Raschka nested cross-validation method was used, which entails having 5 outer folds and 2 inner folds. The 5 models created at the end of the outer loop of cross validation then had their performance averaged to provide the unbiased algorithm performance estimate and the standard deviation of this estimate is calculated as well to provide a measure of variation. Afterward the mean R2 was also calculated for performance evaluation.

#### Final data set creation with missing value imputation using MissForest.
MissForest imputation was selected as the missing value imputation method because it is a robust imputation method that is nonparametric, can handle numeric and categorical variables, is computationally inexpensive, can handle high dimensionality, and provides error estimates using OOB error.

```{r Final Dataset Cleaning and Imputation for Severity}
#Joining together most of the final data set
Dat_sv <- securities2 %>% left_join(stocks2, by = c("gvkey")) #%>% add_row(gvkey = "025119")
Dat_sv <- Dat_sv %>% left_join(fundamentals2, by = c("gvkey"))
Dat_sv <- Dat_sv %>% left_join(ratings2, by = c("gvkey"))
Dat_sv <- Dat_sv %>% left_join(ggroup_df %>% distinct(.keep_all = TRUE), by = c("gvkey"))

#Adding the final variables of interest to the data frame, settlement amount
Dat_sv <- Dat_sv %>% left_join(SCA_filings2, by = c("gvkey"))

#removing Mohawk Industries Observation temporarily
Mohawk_Obs_sv <- Dat_sv %>% filter(gvkey == "025119")
Mohawk_Obs_sv

#filtering the data to only observations with a settlement amount
#since this will be our response variable
Dat_sv <-  Dat_sv %>% filter(SettlementAmount > 0)

#rebinding the Mohawk Industries Observation so it gets missing values imputed
Dat_sv <- rbind(Dat_sv, Mohawk_Obs_sv)

Dat_sv <- Dat_sv %>%
  #removing variables that are no longer relevant, only used as foreign keys
  dplyr::select(-c(FilingYear, Sued, Ticker)) %>%
  #also only using companies that we have some info about (have a gvkey identified)
  filter(!is.na(gvkey))

#checking NAs, quite a lot
NAcheck(Dat_sv) %>% arrange(percent_of_missing_values)
#removing columns with too many NAs
Dat_sv <- NAcull(Dat_sv)

#visualizing missing data to help with MAR assessment
#aggr_plot <- aggr(Dat_sv, col=c('navyblue','red'), numbers=TRUE, sortVars=TRUE, labels=names(data), cex.axis=.7, gap=3, ylab=c("Histogram of missing #data","Pattern"))

#Saving settlement data so it can be removed and reattached
#so it isn't imputed
Dat_sv_Settle <- Dat_sv %>% dplyr::select(c(SettlementAmount, gvkey))
Dat_sv <- Dat_sv %>% dplyr::select(-any_of(c("SettlementAmount", "gvkey", "gsector")))
Dat_sv <- as.data.frame(Dat_sv)

#performing missForest imputation
#https://pubmed.ncbi.nlm.nih.gov/22039212/
#a good non parametric NA solution
set.seed(123, kind = "L'Ecuyer-CMRG")
Dat_sv_imputed <- missForest(Dat_sv, parallelize = "forests", verbose = TRUE, maxiter = 3, ntree = 200)
Dat_sv_imputed$OOBerror

#checking the imputed data
Dat_sv_best_impute <- Dat_sv_imputed$ximp
NAcheck(Dat_sv_best_impute)

#reattaching settlement amount data after missing value imputation
Dat_sv <- cbind(Dat_sv_Settle, Dat_sv_best_impute)

#saving Mohawk industries observation for future prediction
Mohawk_Obs_sv <- Dat_sv %>% filter(gvkey == "025119")
Mohawk_Obs_sv

#removing Mohawk Industries Observation so it isn't used in either
#the training or testing sets
Dat_sv <- Dat_sv %>% filter(gvkey != "025119")

#capping outliers
Dat_sv$SettlementAmount <- capOutlier(Dat_sv$SettlementAmount)

#assessing the data
skimr::skim(Dat_sv)
```

#### The Boruta Algorithm was used for feature selection to reduce dimensionality of the data for the small severity data set.

```{r Boruta Feature Selection for Severity}
#for reproducibility
set.seed(111)

#I use Boruta, a wrapper method that selects all relevant variables for feature selection
boruta_sv <- Boruta(SettlementAmount ~ ., data = Dat_sv[,-2], doTrace = 2, maxRuns = 500)
print(boruta_sv)

#the default plot of boruta, a bit too dense to be that informative besides
#the fact it is working
plot(boruta_sv, las = 2, cex.axis = 0.7)
attStats(boruta_sv) %>% arrange(desc(meanImp))

#tentative rough fix fixes features that are still tenative after the maxRuns is reached
#for the Boruta feature selection algorithm
final_boruta_sv <- TentativeRoughFix(boruta_sv)
features_selected_sv <- getSelectedAttributes(final_boruta_sv)
features_selected_sv <- c(features_selected_sv, "SettlementAmount", "gvkey")

#plotting the top 10% of variables by mean importance in regards to being sued
attStats(boruta_sv) %>%
  mutate(Variable = rownames(attStats(boruta_sv))) %>% 
    dplyr::filter(meanImp > quantile(attStats(boruta_sv)$meanImp, .95)) %>%
   mutate(Boruta_Mean_Importance = meanImp,
    Variable = fct_reorder(Variable, meanImp)) %>%
  ggplot(aes(x = Boruta_Mean_Importance, y = Variable)) +
  geom_col(fill = "tomato") +
  labs(title = "Top 5% Most Important Variable for SettlementAmount as Response", y = NULL)
```

#### A tidymodels recipe was used to remove all zero variance variables, normalize the data, and one hot encode all categorical variables.

```{r Severity Recipes Feature Engineering}
set.seed(3)

#Generic Recipe for Feature Engineering
Dat_sv_rec <- recipe(SettlementAmount ~ ., data = Dat_sv[,features_selected_sv]) %>%
  update_role(gvkey, new_role = "ID") %>%
  update_role(SettlementAmount, new_role = "outcome") %>%
  #removing all zero variance variables
  step_zv() %>%
  #normalization so feature scales are commensurate
  step_normalize(all_numeric(), -c(gvkey, SettlementAmount)) %>%
  #turning nominal factors into dummy variables so they are compatible with certain modeling functions
  step_dummy(all_nominal(), -gvkey)

#code to assess if the recipe is working properly
#test2 <- 
#  Dat_sv_rec %>% # use the recipe object
#  prep() %>% # perform the recipe on training data
#  juice() #returns the data with all recipe steps applied

#creating a workflow with the data preprocessing recipe
#will be used when tuning each model
Dat_sv_wf <- workflow() %>%
  add_recipe(Dat_sv_rec)

#histogram of settlement amounts
ggplot(Dat_sv, aes(SettlementAmount)) +
  geom_histogram(color = "#000000", fill = "#85C285", bins = 20) +
  #geom_text(data = annotations, aes(x = x, y = y, label = paste(label, x)), size = 5, fontface = "bold") +
  scale_x_continuous(labels=scales::dollar_format()) +
  labs(title = "Histogram of Sharehold Settlement Amounts", x = "Settlement Amount", y = "Count") +
  theme_classic()
```

#### The metrics of algorithm evaluation were RMSE and R-squared.

```{r Severity Train Test Split}
#number of outer folds for the nested cross validation procedure
outer_folds_k <- 5

#creating the outer folds
set.seed(4)
outer_folds_i <- sample(rep(1:outer_folds_k, length.out = nrow(Dat_sv)))

#setting the evaluation metrics
metrics_sv <- metric_set(yardstick::rmse, yardstick::rsq)
```

#### Performing nested cross validation for algorithm selection.

```{r Nested Cross Validation Severity RandomForest}
#random forest tuneable regression model specifications
rf_sv_spec <- 
  rand_forest(mtry = tune(), min_n = tune(), trees = 1000) %>% 
  set_engine("ranger", num.threads = parallel::detectCores(), importance = "impurity") %>% 
  set_mode("regression")

#creating a tibble to store the metrics of the best hyperparameter optimized
#models of each loop
best_inner_error <- tibble(.metric = character(), 
                           .estimate = double())

#for loop that goes through each outer fold
for (k in 1:outer_folds_k) {
  
  #each outer fold has a chance to be the test set
  #while the rest are designated training data
  test_i <- which(outer_folds_i == k)
  train <- Dat_sv[-test_i, ]
  train <- train[,features_selected_sv]
  test <- Dat_sv[test_i, ]
  test <- test[,features_selected_sv]
  
  #work around to turn the manual data split into a rsplit object
  #for use with tidymodels
  df <- bind_rows(train, test, .id = "dataset") %>% 
    mutate(dataset = factor(dataset, labels = c("train", "test")))
  train_ids <- which(df$dataset == "train")
  outer_rsplit <- initial_split(df)
  outer_rsplit$in_id <- train_ids
  
  #create 2 inner folds from the data designated as the training data for this loop
  #we're using the Raschka method of nested CV
  set.seed(5)
  Dat_sv_inner_folds <- vfold_cv(train, v = 2)
  
  #optimize hyper-parameters using these 2 inner folds
  rf_sv_tune_results <- Dat_sv_wf %>%
  add_model(rf_sv_spec) %>% 
  tune_grid(Dat_sv_inner_folds,
            grid = 20, #20 candidate models
            control = control_grid(save_pred = TRUE),
            metrics = metrics_sv)
  
  #checking notes to debug some errors
  rf_sv_tune_results$.notes
  
  #store the best model
  rf_sv_best <- rf_sv_tune_results %>% 
    select_best("rmse")
  
  #create a work flow using the best model
  rf_sv_wf <- Dat_sv_wf %>% add_model(rf_sv_spec)
  best_rf_sv_wf <- finalize_workflow(rf_sv_wf, rf_sv_best)
  
  #train and evaluate the best model on the training set
  best_sv_rf_metrics <- last_fit(best_rf_sv_wf, outer_rsplit, metrics = metrics_sv) %>% 
    collect_metrics() %>% 
    dplyr::select(-c(".estimator", ".config"))
  
  #append these metrics to a tibble
  #in the end will have 5 best models that will be averaged
  #to estimate performance of the algorithm
  best_inner_error <- best_inner_error %>% add_row(best_sv_rf_metrics)
}

#saving the random forest performance data 
rf_best_inner_error <- best_inner_error
rf_outer_error_mean <- rf_best_inner_error %>% filter(.metric == "rmse") %>% dplyr::select(.estimate) %>% pull() %>% mean()
rf_outer_error_std <- rf_best_inner_error %>% filter(.metric == "rmse") %>% dplyr::select(.estimate) %>% pull() %>% std_mean()
rf_outer_error_r2_mean <- rf_best_inner_error %>% filter(.metric == "rsq") %>% dplyr::select(.estimate) %>% pull() %>% na_rm_mean()

#outputting the mean R-squared for a rough performance estimate
rf_outer_error_r2_mean
```

```{r Nested Cross Validation Severity Lasso}
#lasso regression model specifications
#mixture 1 indicates this is a pure lasso model
#thus variable selection is performed (via shrinking coef to 0)
lassor_spec <- linear_reg(penalty = tune(), mixture = 1) %>%
  set_engine("glmnet")

#grid for tuning the penalty of the lasso regression model
lassor_grid <- tibble(penalty = 10^seq(8, -8, by = -.25))

#creating a tibble to store the metrics of the best hyperparameter optimized
#models of each loop
best_inner_error <- tibble(.metric = character(), 
                           .estimate = double())

#for loop that goes through each outer fold
for (k in 1:outer_folds_k) {
  
  #each outer fold has a chance to be the test set
  #while the rest are designated training data
  test_i <- which(outer_folds_i == k)
  train <- Dat_sv[-test_i, ]
  train <- train[,features_selected_sv]
  test <- Dat_sv[test_i, ]
  test <- test[,features_selected_sv]
  
  #work around to turn the manual data split into a rsplit object
  #for use with tidymodels
  df <- bind_rows(train, test, .id = "dataset") %>% 
    mutate(dataset = factor(dataset, labels = c("train", "test")))
  train_ids <- which(df$dataset == "train")
  outer_rsplit <- initial_split(df)
  outer_rsplit$in_id <- train_ids
  
  #create 2 inner folds from the data designated as the training data for this loop
  #we're using the Raschka method of nested CV
  set.seed(33)
  Dat_sv_inner_folds <- vfold_cv(train, v = 2)
  
  #optimize hyper-parameters using these 2 inner folds
  lassor_sv_tune_results <- Dat_sv_wf %>%
    add_model(lassor_spec) %>% 
    tune_grid(Dat_sv_inner_folds,
            grid = lassor_grid,
            control = control_grid(save_pred = TRUE),
            metrics = metrics_sv)
  
  #store the best model
  lassor_sv_best <- lassor_sv_tune_results %>% 
    select_best("rmse")
  
  #create a work flow using the best model
  lassor_sv_wf <- Dat_sv_wf %>% add_model(lassor_spec)
  best_lassor_sv_wf <- finalize_workflow(lassor_sv_wf, lassor_sv_best)
  
  #train and evaluate the best model on the training set
  best_sv_lassor_metrics <- last_fit(best_lassor_sv_wf, outer_rsplit, metrics = metrics_sv) %>% 
    collect_metrics() %>% 
    dplyr::select(-c(".estimator", ".config"))
  
  #append these metrics to a tibble
  #in the end will have 5 best models that will be averaged
  #to estimate performance of the algorithm
  best_inner_error <- best_inner_error %>% add_row(best_sv_lassor_metrics)
}

#saving the lasso regression performance data 
lassor_best_inner_error <- best_inner_error
lassor_outer_error_mean <- lassor_best_inner_error %>% filter(.metric == "rmse") %>% dplyr::select(.estimate) %>% pull() %>% mean()
lassor_outer_error_std <- lassor_best_inner_error %>% filter(.metric == "rmse") %>% dplyr::select(.estimate) %>% pull() %>% std_mean()
lassor_outer_error_r2_mean <- lassor_best_inner_error %>% filter(.metric == "rsq") %>% dplyr::select(.estimate) %>% pull() %>% na_rm_mean()

#outputting the mean R-squared for a rough performance estimate
lassor_outer_error_r2_mean
```

```{r Nested Cross Validation Severity Ridge}
#lasso regression model specifications
#mixture 0 indicates this is a pure lasso model
#thus variable selection is performed (via shrinking coef to 0)
ridger_spec <- linear_reg(penalty = tune(), mixture = 0) %>%
  set_engine("glmnet")

#grid for tuning the penalty of the lasso regression model
ridger_grid <- tibble(penalty = 10^seq(8, -8, by = -.25))

#creating a tibble to store the metrics of the best hyperparameter optimized
#models of each loop
best_inner_error <- tibble(.metric = character(), 
                           .estimate = double())

#for loop that goes through each outer fold
for (k in 1:outer_folds_k) {
  
  #each outer fold has a chance to be the test set
  #while the rest are designated training data
  test_i <- which(outer_folds_i == k)
  train <- Dat_sv[-test_i, ]
  train <- train[,features_selected_sv]
  test <- Dat_sv[test_i, ]
  test <- test[,features_selected_sv]
  
  #work around to turn the manual data split into a rsplit object
  #for use with tidymodels
  df <- bind_rows(train, test, .id = "dataset") %>% 
    mutate(dataset = factor(dataset, labels = c("train", "test")))
  train_ids <- which(df$dataset == "train")
  outer_rsplit <- initial_split(df)
  outer_rsplit$in_id <- train_ids
  
  #create 2 inner folds from the data designated as the training data for this loop
  #we're using the Raschka method of nested CV
  set.seed(33)
  Dat_sv_inner_folds <- vfold_cv(train, v = 2)
  
  #optimize hyper-parameters using these 2 inner folds
  ridger_sv_tune_results <- Dat_sv_wf %>%
    add_model(ridger_spec) %>% 
    tune_grid(Dat_sv_inner_folds,
            grid = ridger_grid,
            control = control_grid(save_pred = TRUE),
            metrics = metrics_sv)
  
  #store the best model
  ridger_sv_best <- ridger_sv_tune_results %>% 
    select_best("rmse")
  
  #create a work flow using the best model
  ridger_sv_wf <- Dat_sv_wf %>% add_model(ridger_spec)
  best_ridger_sv_wf <- finalize_workflow(ridger_sv_wf, ridger_sv_best)
  
  #train and evaluate the best model on the training set
  best_sv_ridger_metrics <- last_fit(best_ridger_sv_wf, outer_rsplit, metrics = metrics_sv) %>% 
    collect_metrics() %>% 
    dplyr::select(-c(".estimator", ".config"))
  
  #append these metrics to a tibble
  #in the end will have 5 best models that will be averaged
  #to estimate performance of the algorithm
  best_inner_error <- best_inner_error %>% add_row(best_sv_ridger_metrics)
}

#saving the ridge regression performance data 
ridger_best_inner_error <- best_inner_error
ridger_outer_error_mean <- ridger_best_inner_error %>% filter(.metric == "rmse") %>% dplyr::select(.estimate) %>% pull() %>% mean()
ridger_outer_error_std <- ridger_best_inner_error %>% filter(.metric == "rmse") %>% dplyr::select(.estimate) %>% pull() %>% std_mean()
ridger_outer_error_r2_mean <- ridger_best_inner_error %>% filter(.metric == "rsq") %>% dplyr::select(.estimate) %>% pull() %>% mean()

#outputting the mean R-squared for a rough performance estimate
ridger_outer_error_r2_mean
```

```{r Nested Cross Validation Severity XGBoost}
#xgb tuneable model specifications
xgb_sv_spec <- boost_tree(
  trees = 1000, 
  tree_depth = tune(),
  min_n = tune(), 
  loss_reduction = tune(),                     
  sample_size = tune(), mtry = tune(), ## randomness
  learn_rate = tune(), ## step size
  ) %>% 
  set_engine("xgboost") %>% 
  set_mode("regression")

#for tuning, tries to cover parameter space efficiently
xgb_sv_grid <- grid_latin_hypercube(
  tree_depth(),
  min_n(),
  loss_reduction(),
  sample_size = sample_prop(),
  finalize(mtry(), Data_train),
  learn_rate(),
  size = 40
)

#creating a tibble to store the metrics of the best hyperparameter optimized
#models of each loop
best_inner_error <- tibble(.metric = character(), 
                           .estimate = double())

#for loop that goes through each outer fold
for (k in 1:outer_folds_k) {
  
  #each outer fold has a chance to be the test set
  #while the rest are designated training data
  test_i <- which(outer_folds_i == k)
  train <- Dat_sv[-test_i, ]
  train <- train[,features_selected_sv]
  test <- Dat_sv[test_i, ]
  test <- test[,features_selected_sv]
  
  #work around to turn the manual data split into a rsplit object
  #for use with tidymodels
  df <- bind_rows(train, test, .id = "dataset") %>% 
    mutate(dataset = factor(dataset, labels = c("train", "test")))
  train_ids <- which(df$dataset == "train")
  outer_rsplit <- initial_split(df)
  outer_rsplit$in_id <- train_ids
  
  #create 2 inner folds 
  #from the data designated as the training data for this loop
  #we're using the Raschka method of nested CV (5 outer folds, 2 inner folds)
  set.seed(33)
  Dat_sv_inner_folds <- vfold_cv(train, v = 2)
  
  #optimize hyper-parameters using these 2 inner folds
  XGBoost_sv_tune_results <- Dat_sv_wf %>%
    add_model(xgb_sv_spec) %>% 
    tune_grid(Dat_sv_inner_folds,
            grid = xgb_sv_grid,
            control = control_grid(save_pred = TRUE),
            metrics = metrics_sv)
  
  #store the best model
  XGBoost_sv_best <- XGBoost_sv_tune_results %>% 
    select_best("rmse")
  
  #create a work flow using the best model
  XGBoost_sv_wf <- Dat_sv_wf %>% add_model(xgb_sv_spec)
  best_XGBoost_sv_wf <- finalize_workflow(XGBoost_sv_wf, XGBoost_sv_best)
  
  #train and evaluate the best model on the training set
  best_sv_XGBoost_metrics <- last_fit(best_XGBoost_sv_wf, outer_rsplit, metrics = metrics_sv) %>% 
    collect_metrics() %>% 
    dplyr::select(-c(".estimator", ".config"))
  
  #append these metrics to a tibble
  #in the end will have 5 best models that will be averaged
  #to estimate performance of the algorithm
  best_inner_error <- best_inner_error %>% add_row(best_sv_XGBoost_metrics)
  print(paste("Run complete", k))
}

#saving the XGBoost performance data 
XGBoost_best_inner_error <- best_inner_error
XGBoost_outer_error_mean <- XGBoost_best_inner_error %>% filter(.metric == "rmse") %>% dplyr::select(.estimate) %>% pull() %>% mean()
XGBoost_outer_error_std <- XGBoost_best_inner_error %>% filter(.metric == "rmse") %>% dplyr::select(.estimate) %>% pull() %>% std_mean()
XGBoost_outer_error_r2_mean <- XGBoost_best_inner_error %>% filter(.metric == "rsq") %>% dplyr::select(.estimate) %>% pull() %>% na_rm_mean()

#outputting the mean R-squared for a rough performance estimate
XGBoost_outer_error_r2_mean
```

#### Comparing the mean outer fold performance of each algorithm (model building procedure).

```{r Severity Nested CV Comparison}

#creating a data frame from all those performance data sets saved at the end of each
#nested cross validation procedure for each algorithm
Sv_mdl_outer_errors <- bind_rows(tibble(rmse = rf_outer_error_mean, std = rf_outer_error_std, r2 = rf_outer_error_r2_mean, 
                                        model = "Random Forest"),
                               tibble(rmse = lassor_outer_error_mean, std = lassor_outer_error_std, r2 = lassor_outer_error_r2_mean,
                                      model = "Lasso Regression"),
                               tibble(rmse = ridger_outer_error_mean, std = ridger_outer_error_std, r2 = ridger_outer_error_r2_mean,
                                      model = "Ridge Regression"),
                               tibble(rmse = XGBoost_outer_error_mean, std = XGBoost_outer_error_std, r2 = XGBoost_outer_error_r2_mean,
                                      model = "XGBoost"))

#making bar plots of the average outer fold RMSE of each algorithm
Sv_mdl_outer_errors %>% 
  arrange(rmse, decreasing = TRUE) %>%
  mutate(model = fct_reorder(model, rmse, .desc = TRUE)) %>%
  ggplot(aes(model, rmse, fill=model)) +
  geom_col(show.legend = FALSE) +
  geom_errorbar(aes(ymin=rmse-(std/sqrt(5)*1.96), ymax=rmse+(std/sqrt(5)*1.96)), width= .3,
                 position=position_dodge(.9)) +
  scale_fill_manual(values = c("Random Forest"="#00BFC4",
                               "Lasso Regression"="#FC717F",
                               "Ridge Regression"="#E7861B",
                               "XGBoost"="#C77CFF")) +
  coord_flip() +
  geom_text(size = 4, aes(label = round(rmse, 3), y = rmse), vjust = 2.5) +
  labs(title = "Best Severity Model by Mean Outer Test Fold RMSE", x = "Model", y = "RMSE")

#making bar plots of the average outer fold R-squared of each algorithm
Sv_mdl_outer_errors %>% 
  arrange(r2, decreasing = TRUE) %>%
  mutate(model = fct_reorder(model, r2, .desc = FALSE)) %>%
  ggplot(aes(model, r2, fill=model)) +
  geom_col(show.legend = FALSE) +
  scale_fill_manual(values = c("Random Forest"="#00BFC4",
                               "Lasso Regression"="#FC717F",
                               "Ridge Regression"="#E7861B",
                               "XGBoost"="#C77CFF")) +
  coord_flip() +
  geom_text(size = 4, aes(label = round(r2, 3), y = r2 - .05), vjust = 1) +
  labs(title = "Average Outer Test Fold R-Squared", x = "Model", y = "R-Squared")
```

#### Finally, constructing a prediction interval of the cost of potential SCA litigation for Mohawk Industries with the best algorithm.

```{r Best Severity Prediction}
#now according to Raschka we refit the best algorithm on the whole data set
#with hyper parameters optimized on cross validation using the whole data set
#and predict 

#creating folds
Dat_sv_folds <- vfold_cv(train, v = 5)

#final tuning using cross validation
final_rf_sv_tune_results <- Dat_sv_wf %>%
  add_model(rf_sv_spec) %>% 
  tune_grid(Dat_sv_folds,
            grid = 20, #20 candidate models
            control = control_grid(save_pred = TRUE),
            metrics = metrics_sv)

#finalizing work flow
final_best_rf_sv_wf <- finalize_workflow(rf_sv_wf, final_rf_sv_tune_results %>% select_best("rmse"))

#fitting the final model with optimized hyper-parameters
final_fit_best_rf_sv_wf <- fit(final_best_rf_sv_wf, data = Dat_sv[,features_selected_sv])

#make a nice feature importance plot
final_fit_best_rf_sv_wf %>%
extract_fit_parsnip() %>% 
  vip(num_features = 10)

#final point prediction
stats::predict(final_fit_best_rf_sv_wf, Mohawk_Obs_sv[,features_selected_sv])

#To construct a confidence interval we'll use the naive MAPIE method
pred_quantiles <- final_rf_sv_tune_results %>% collect_predictions() %>% mutate(residual = abs(SettlementAmount-.pred)) %>% dplyr::select(residual) %>% pull() %>% quantile(probs = c(.9, .95))

#using the 95th quantile of absolute error to construct a 95% prediction interval
#of the severity estimate of SCA litigation for Mohawk Industries Inc.
bind_rows(Lower = predict(final_fit_best_rf_sv_wf, Mohawk_Obs_sv[,features_selected_sv]) - pred_quantiles[2], Upper = predict(final_fit_best_rf_sv_wf, Mohawk_Obs_sv[,features_selected_sv]) + pred_quantiles[2])

#plotting the absolute errors
#with a red line to indicate the .95 quantile
final_rf_sv_tune_results %>% collect_predictions() %>% mutate(residual = abs(SettlementAmount-.pred)) %>%
  ggplot(aes(residual)) +
  geom_histogram(color = "#000000", fill = "#85C285", bins = 20) +
  #geom_text(data = annotations, aes(x = x, y = y, label = paste(label, x)), size = 5, fontface = "bold") +
  scale_x_continuous(labels=scales::dollar_format()) +
  geom_vline(xintercept= pred_quantiles[2], 
             linetype="dashed", 
             color = "red",
             size = .5) +
  labs(title = "Absolute Calibration Errors for the Final Random Forest Model", caption = "Red line indicates .95 quantile") +
  theme_classic()
```

```{r extraneous lasso tune visual code, eval = FALSE}
# ridger_tune_results %>%
#   collect_metrics() %>%
#   ggplot(aes(penalty, mean, color = .metric)) +
#   geom_errorbar(aes(
#     ymin = mean - std_err,
#     ymax = mean + std_err),
#   alpha = 0.5) +
#   geom_line(size = 1.5) +
#   facet_wrap(~.metric, scales = "free", nrow = 2) +
#   scale_x_log10() +
#   theme(legend.position = "none")
```

```{r MICE code, eval = FALSE}
# #MICE to replace missing values
# init = mice(Data, maxit = 0)
# meth = init$method
# predM = init$predictorMatrix
# 
# #checking methods for variables that are still missing a method
# #after automatic assignment
# meth
# 
# #using cart method
# #also not imputing the variables of interest
# meth[] <- "cart"
# 
# #manually setting method
# meth[names(meth) %in% c("stko", "auop", "ceoso", "cfoso")] <- "polyreg"
# meth[names(meth) %in% c("idbflag")] <- "logreg"
# meth[names(meth) %in% c("SettlementAmount", "Sued")] <- ""
# 
# #holding out the variables of interest as predictors in MICE
# predM[, c("SettlementAmount", "Sued")] <- 0
# 
# #calculating optimal # of imputations
# perc_miss <- NAcheck(Data)
# perc_miss <- mean(perc_miss$percent_of_missing_values)*100
# perc_miss <- perc_miss %/% 1
# 
# #Finally multiple imputation
# set.seed(3)
# temp_imputed = mice(Data, method = meth, maxit = 25, predictorMatrix = predM, m = 5)
# imputed <- complete(temp_imputed)
# 
# #plot(temp_imputed)
# Data <- imputed
# NAcheck(Data)
```
